---
title: "Brining it All Together (Extended Networks Towards Thesis"
description: |
  Here I extend my inital networks from 1000-1300 to the year 1900.
author:
  - name: Noah Milstein
    url: {}
date: 2022-05-03
output:
  distill::distill_article:
    self_contained: false
---

### Loading Required Packages

```{r}

library(readxl)

library(igraph)

library(statnet)

library(network)

library(tidyverse)

library(ggnetwork)

library(GGally)

library(ggplot2)

library(sna)

library(intergraph)

library(ggbiplot)

library(knitr)

library(GGally)

library(aRtsy)

library(stringr)

library(dplyr)

```


#### Loading Datasets

As will be the case in nearly all chunks that follow, I use the same basic, repeated code for each of the 8 datasets from 1000 to 1800. 

```{r setup, include=FALSE}

# First using the read_excel function from readxl I load the excel spreadsheet

wars_in_1000s <- read_excel("~/Desktop/Spring 2022/Networks/wars_in_1000s.xlsx")

# Next I load my in list and out list removing line breaks

wars_in_1000s$"Out-list" <- sapply(wars_in_1000s$"Out-list",
                                    function(x) { gsub("[\r\n]", "", x) })
wars_in_1000s$"In-list"<- sapply(wars_in_1000s$"In-list",
                                    function(x) { gsub("[\r\n]", "", x) })

# This procedure is repeated 8 times.

wars_in_1100s <- read_excel("~/Desktop/Spring 2022/Networks/wars_in_1100s.xlsx")

wars_in_1100s$"Out" <- sapply(wars_in_1100s$"Out",
                                    function(x) { gsub("[\r\n]", "", x) })
wars_in_1100s$"In"<- sapply(wars_in_1100s$"In",
                                    function(x) { gsub("[\r\n]", "", x) })

wars_in_1200s <- read_excel("~/Desktop/Spring 2022/Networks/wars_in_1200s.xlsx")


wars_in_1200s$"Out" <- sapply(wars_in_1200s$"Out",
                                    function(x) { gsub("[\r\n]", "", x) })

wars_in_1200s$"In"<- sapply(wars_in_1200s$"In",  
                                    function(x) { gsub("[\r\n]", "", x) })

wars_in_1300s <- read_excel("~/Desktop/Spring 2022/Networks/wars_in_1300s.xlsx")

wars_in_1300s$"In" <- sapply(wars_in_1300s$"Out-List",
                                    function(x) { gsub("[\r\n]", "", x) })
wars_in_1300s$"Out"<- sapply(wars_in_1300s$"In-List",  
                                    function(x) { gsub("[\r\n]", "", x) })

wars_in_1300s <- wars_in_1300s[,c(6:7)]

# I then had non-breaking spaces so I had to use str_trim()

wars_in_1300s$In <- str_trim(wars_in_1300s$In)

wars_in_1300s$Out <- str_trim(wars_in_1300s$Out)

wars_in_1400s <- read_excel("~/Desktop/Spring 2022/Networks/wars_in_1400s.xlsx")

wars_in_1400s$"In" <- sapply(wars_in_1400s$"In-List",
                                    function(x) { gsub("[\r\n]", "", x) })
wars_in_1400s$"Out"<- sapply(wars_in_1400s$"Out-List",  
                                    function(x) { gsub("[\r\n]", "", x) })

wars_in_1400s <- wars_in_1400s[,c(6:7)]

wars_in_1400s$In <- str_trim(wars_in_1400s$In)

wars_in_1400s$Out <- str_trim(wars_in_1400s$Out)

wars_in_1500s <- read_excel("~/Desktop/Spring 2022/Networks/wars_in_1500s.xlsx")

wars_in_1500s$"In" <- sapply(wars_in_1500s$"In-List",
                                    function(x) { gsub("[\r\n]", "", x) })
wars_in_1500s$"Out"<- sapply(wars_in_1500s$"Out-List",  
                                    function(x) { gsub("[\r\n]", "", x) })

wars_in_1500s <- wars_in_1500s[,c(6:7)]

wars_in_1500s$In <- str_trim(wars_in_1500s$In)

wars_in_1500s$Out <- str_trim(wars_in_1500s$Out)

wars_in_1600s <- read_excel("~/Desktop/Spring 2022/Networks/wars_in_1600s.xlsx")

wars_in_1600s$"In" <- sapply(wars_in_1600s$"In-List",
                                    function(x) { gsub("[\r\n]", "", x) })
wars_in_1600s$"Out"<- sapply(wars_in_1600s$"Out-List",  
                                    function(x) { gsub("[\r\n]", "", x) })

wars_in_1600s <- wars_in_1600s[,c(6:7)]

wars_in_1600s$In <- str_trim(wars_in_1600s$In)

wars_in_1600s$Out <- str_trim(wars_in_1600s$Out)

wars_in_1700s <- read_excel("~/Desktop/Spring 2022/Networks/wars_in_1700s.xlsx")

wars_in_1700s$"In" <- sapply(wars_in_1700s$"In-List",
                                    function(x) { gsub("[\r\n]", "", x) })
wars_in_1700s$"Out"<- sapply(wars_in_1700s$"Out-List",  
                                    function(x) { gsub("[\r\n]", "", x) })

wars_in_1700s <- wars_in_1700s[,c(6:7)]

wars_in_1700s$In <- str_trim(wars_in_1700s$In)

wars_in_1700s$Out <- str_trim(wars_in_1700s$Out)

wars_in_1800s <- read_excel("~/Desktop/Spring 2022/Networks/wars_in_1800s.xlsx")

wars_in_1800s$In <- sapply(wars_in_1800s$"In-List",
                                    function(x) { gsub("[\r\n]", "", x) })
wars_in_1800s$Out<- sapply(wars_in_1800s$"Out-List",  
                                    function(x) { gsub("[\r\n]", "", x) })

wars_in_1800s <- wars_in_1800s[,c(6:7)]

wars_in_1800s$In <- str_trim(wars_in_1800s$In)

wars_in_1800s$Out <- str_trim(wars_in_1800s$Out)

```

#### Creating Networks

Subsequently I must make my excel spreadsheet into a variety of network types for later analysis.

```{r}

# First I used as.matrix from base to make the data frame into a matrix.

wars_in_1000s_edgelist <- as.matrix(wars_in_1000s)

# Next I make a igraph object first using graph.edgelist from igraph

wars_in_1000s_edgelist_network_edgelist <- 
  
  graph.edgelist(wars_in_1000s_edgelist, directed=TRUE)

# To Simplify this I then just use graph_from_data_frame() from Igraph

wars_in_1000s.ig<-graph_from_data_frame(wars_in_1000s)

# In order to make a network obejct from statnet and sna I then use 
# asNetwork from intergraph to coerce the igraph network to a statnet object.

wars_in_1000s_network <- asNetwork(wars_in_1000s.ig)

# I finally convert the igraph object back to a matrix to make sure information

# was not lost during the conversion. This process is repeated.

wars_1000s<-as.matrix(as_adjacency_matrix(wars_in_1000s.ig))

wars_in_1100s_edgelist <- as.matrix(wars_in_1100s)

wars_in_1100s_edgelist_network_edgelist <- 
  
  graph.edgelist(wars_in_1100s_edgelist, directed=TRUE)

wars_in_1100s.ig<-graph_from_data_frame(wars_in_1100s)

wars_in_1100s_network <- asNetwork(wars_in_1100s.ig)

wars_1100s<-as.matrix(as_adjacency_matrix(wars_in_1100s.ig))

wars_in_1200s_edgelist <- as.matrix(wars_in_1200s)

wars_in_1200s_edgelist_network_edgelist <- 
  
  graph.edgelist(wars_in_1200s_edgelist, directed=TRUE)

wars_in_1200s.ig <- graph_from_data_frame(wars_in_1200s)

wars_in_1200s_network <- asNetwork(wars_in_1200s.ig)

wars_1200s <- as.matrix(as_adjacency_matrix(wars_in_1200s.ig))

wars_in_1300s_edgelist <- as.matrix(wars_in_1300s)

wars_in_1300s_edgelist_network_edgelist <- 
  
  graph.edgelist(wars_in_1300s_edgelist, directed=TRUE)

wars_in_1300s.ig<-graph_from_data_frame(wars_in_1300s)

wars_in_1300s_network <- asNetwork(wars_in_1300s.ig)

wars_1300s <- as.matrix(as_adjacency_matrix(wars_in_1300s.ig))

wars_in_1400s_edgelist <- as.matrix(wars_in_1400s)

wars_in_1400s_edgelist_network_edgelist <- 
  
  graph.edgelist(wars_in_1400s_edgelist, directed=TRUE)

wars_in_1400s.ig<-graph_from_data_frame(wars_in_1400s)

wars_in_1400s_network <- asNetwork(wars_in_1400s.ig)

wars_1400s<-as.matrix(as_adjacency_matrix(wars_in_1400s.ig))

wars_in_1500s_edgelist <- as.matrix(wars_in_1500s)

wars_in_1500s_edgelist_network_edgelist <- 
  
  graph.edgelist(wars_in_1500s_edgelist, directed=TRUE)

wars_in_1500s.ig<-graph_from_data_frame(wars_in_1500s)

wars_in_1500s_network <- asNetwork(wars_in_1500s.ig)

wars_1500s<-as.matrix(as_adjacency_matrix(wars_in_1500s.ig))

wars_in_1600s_edgelist <- as.matrix(wars_in_1600s)

wars_in_1600s_edgelist_network_edgelist <- 
  
  graph.edgelist(wars_in_1600s_edgelist, directed=TRUE)

wars_in_1600s.ig <- graph_from_data_frame(wars_in_1600s)

wars_in_1600s_network <- asNetwork(wars_in_1600s.ig)

wars_1600s <- as.matrix(as_adjacency_matrix(wars_in_1600s.ig))

wars_in_1700s_edgelist <- as.matrix(wars_in_1700s)

wars_in_1700s_edgelist_network_edgelist <- 
  
  graph.edgelist(wars_in_1700s_edgelist, directed=TRUE)

wars_in_1700s.ig<-graph_from_data_frame(wars_in_1700s)

wars_in_1700s_network <- asNetwork(wars_in_1700s.ig)

wars_1700s<-as.matrix(as_adjacency_matrix(wars_in_1700s.ig))

wars_in_1800s_edgelist <- as.matrix(wars_in_1800s)

wars_in_1800s_edgelist_network_edgelist <- 
  
  graph.edgelist(wars_in_1800s_edgelist, directed=TRUE)

wars_in_1800s.ig<-graph_from_data_frame(wars_in_1800s)

wars_in_1800s_network <- asNetwork(wars_in_1800s.ig)

wars_1800s <- as.matrix(as_adjacency_matrix(wars_in_1800s.ig))

```

### Inital Network Analysis and Statistics

#### Dyad Census

In order to ensure proper loading and formatting we first look at the dyad census of our networks using sna.

```{r, echo=FALSE}

sna::dyad.census(wars_in_1000s_network) %>% kable()

sna::dyad.census(wars_in_1100s_network) %>% kable()

sna::dyad.census(wars_in_1200s_network) %>% kable()

sna::dyad.census(wars_in_1300s_network) %>% kable()

sna::dyad.census(wars_in_1400s_network) %>% kable()

sna::dyad.census(wars_in_1500s_network) %>% kable()

sna::dyad.census(wars_in_1600s_network) %>% kable()

sna::dyad.census(wars_in_1700s_network) %>% kable()

sna::dyad.census(wars_in_1800s_network) %>% kable()

```

#### Triad Census

Following this we look at the triad census in our data. This is conceptually important for conflict prediction going forward. Certain types of triads may be effective at indicating instability between nations and give us some sense of future conflict development along with general network stability.

```{r, echo=FALSE}

sna::triad.census(wars_in_1000s_network, mode="graph") %>% kable()

sna::triad.census(wars_in_1100s_network, mode="graph") %>% kable()

sna::triad.census(wars_in_1200s_network, mode="graph") %>% kable()

sna::triad.census(wars_in_1300s_network, mode="graph") %>% kable()

sna::triad.census(wars_in_1400s_network, mode="graph") %>% kable()

sna::triad.census(wars_in_1500s_network, mode="graph") %>% kable()

sna::triad.census(wars_in_1600s_network, mode="graph") %>% kable()

sna::triad.census(wars_in_1700s_network, mode="graph") %>% kable()

sna::triad.census(wars_in_1800s_network, mode="graph") %>% kable()

```

#### Transitivity Attributes for Each Century

##### Global Transitivity Attributes for Each Century

Looking at this attributes individually is useful, however in order to understand how they change over time we will concatenate the scores and graph them.

```{r}

transitivity_over_time_global <-c(
  
transitivity(wars_in_1000s.ig, type="global"),

transitivity(wars_in_1100s.ig, type="global"),

transitivity(wars_in_1200s.ig, type="global"),

transitivity(wars_in_1300s.ig, type="global"),

transitivity(wars_in_1400s.ig, type="global"),

transitivity(wars_in_1500s.ig, type="global"),

transitivity(wars_in_1600s.ig, type="global"),

transitivity(wars_in_1700s.ig, type="global"),

transitivity(wars_in_1800s.ig, type="global")

)


network_atts_for_plot <- data.frame(transitivity_over_time_global)
network_atts_for_plot

```

##### Century Attributes, Numeric

I then create a list of doubles to indicate my century of conflict in subsequent analysis.

```{r}

century_for_graphing <- c(1000, 1100, 1200, 1300, 1400, 1500, 1600, 1700, 1800)

network_atts_for_plot$century_for_graphing <- century_for_graphing

```

#### Global Transitivity Over Time Graphed

I then use ggplot to plot the network's transitivity over time using global transitivity.

```{r}


ggplot(network_atts_for_plot) +
  
geom_point(
  
  color="darkorange2", 
  
  aes(
    
    x=century_for_graphing, 
                                
      y=transitivity_over_time_global

    )
)


```

##### Average Transitivity Attributes for Each Century


```{r}

transitivity_over_time_average <-c(
  
transitivity(wars_in_1000s.ig, type="average"),

transitivity(wars_in_1100s.ig, type="average"),

transitivity(wars_in_1200s.ig, type="average"),

transitivity(wars_in_1300s.ig, type="average"),

transitivity(wars_in_1400s.ig, type="average"),

transitivity(wars_in_1500s.ig, type="average"),

transitivity(wars_in_1600s.ig, type="average"),

transitivity(wars_in_1700s.ig, type="average"),

transitivity(wars_in_1800s.ig, type="average")

)

network_atts_for_plot$transitivity_over_time_average <- 
  transitivity_over_time_average

```

#### Average Transitivity Over Time Graphed

These clustering coefficients do differ from century to century between average and global clustering.

```{r}


ggplot(network_atts_for_plot) +
  
geom_point(
  
  color="forestgreen", 
  
  aes(
    
    x=century_for_graphing, 
                                
      y=transitivity_over_time_average

    )
)


```

#### Average Path Length 


```{r}

average_path_length_wars <- c(
  
average.path.length(wars_in_1000s.ig,directed=T),

average.path.length(wars_in_1100s.ig,directed=T),

average.path.length(wars_in_1200s.ig,directed=T),

average.path.length(wars_in_1300s.ig,directed=T),

average.path.length(wars_in_1400s.ig,directed=T),

average.path.length(wars_in_1500s.ig,directed=T),

average.path.length(wars_in_1600s.ig,directed=T),

average.path.length(wars_in_1700s.ig,directed=T),

average.path.length(wars_in_1800s.ig,directed=T)

)


network_atts_for_plot$average_path_length_wars <- 
  average_path_length_wars


```


#### Average Transitivity Over Time Graphed

These clustering coefficients do differ from century to century between average and global clustering.

```{r}


ggplot(network_atts_for_plot) +
  
geom_point(
  
  color="darkslateblue", 
  
  aes(
    
    x=century_for_graphing, 
                                
      y=average_path_length_wars

    )
)


```

#### Density Attributes for Each Century

```{r}

densities_over_time <- c(
  
network.density(wars_in_1000s_network),
  
network.density(wars_in_1100s_network), 
  
network.density(wars_in_1200s_network), 
  
network.density(wars_in_1300s_network),
  
network.density(wars_in_1400s_network),

network.density(wars_in_1500s_network),

network.density(wars_in_1600s_network),

network.density(wars_in_1700s_network),

network.density(wars_in_1800s_network)
)

network_atts_for_plot$densities_over_time <- 
  densities_over_time

```

#### Average Density Over Time Graphed

```{r}


ggplot(network_atts_for_plot) +
  
geom_point(
  
  color="firebrick4", 
  
  aes(
    
    x=century_for_graphing, 
                                
      y=densities_over_time

    )
)


```

#### Centralization Attributes for Each Century

```{r}

centralities_in_degree_over_time<- c(
  
centr_degree(wars_in_1000s.ig, loops = FALSE, mode="in")$centralization,

centr_degree(wars_in_1100s.ig, loops = FALSE, mode="in")$centralization, 

centr_degree(wars_in_1200s.ig, loops = FALSE, mode="in")$centralization, 

centr_degree(wars_in_1300s.ig, loops = FALSE, mode="in")$centralization, 

centr_degree(wars_in_1400s.ig, loops = FALSE, mode="in")$centralization, 

centr_degree(wars_in_1500s.ig, loops = FALSE, mode="in")$centralization, 

centr_degree(wars_in_1600s.ig, loops = FALSE, mode="in")$centralization,

centr_degree(wars_in_1700s.ig, loops = FALSE, mode="in")$centralization,

centr_degree(wars_in_1800s.ig, loops = FALSE, mode="in")$centralization

)

network_atts_for_plot$centralities_in_degree_over_time <- 
  centralities_in_degree_over_time

```

#### Average Centrality Over Time Graphed

```{r}


ggplot(network_atts_for_plot) +
  
geom_point(
  
  color="firebrick4", 
  
  aes(
    
    x=century_for_graphing, 
                                
      y=centralities_in_degree_over_time

    )
)


```

### Beginning with Individual Node Attributes 

#### Nodes Attributes for Each Century Total Degree and Names

```{r}

nodes_1000s <-data.frame(name=wars_in_1000s_network%v%"vertex.names", 
                         degree=sna::degree(wars_in_1000s_network))

nodes_1100s <-data.frame(name=wars_in_1100s_network%v%"vertex.names", 
                         degree=sna::degree(wars_in_1100s_network))

nodes_1200s <-data.frame(name=wars_in_1200s_network%v%"vertex.names", 
                         degree=sna::degree(wars_in_1200s_network))

nodes_1300s <-data.frame(name=wars_in_1300s_network%v%"vertex.names", 
                         degree=sna::degree(wars_in_1300s_network))

nodes_1400s <-data.frame(name=wars_in_1400s_network%v%"vertex.names", 
                         degree=sna::degree(wars_in_1400s_network))

nodes_1500s <-data.frame(name=wars_in_1500s_network%v%"vertex.names", 
                         degree=sna::degree(wars_in_1500s_network))

nodes_1600s <-data.frame(name=wars_in_1600s_network%v%"vertex.names", 
                         degree=sna::degree(wars_in_1600s_network))

nodes_1700s <-data.frame(name=wars_in_1700s_network%v%"vertex.names", 
                         degree=sna::degree(wars_in_1700s_network))

nodes_1800s <-data.frame(name=wars_in_1800s_network%v%"vertex.names", 
                         degree=sna::degree(wars_in_1800s_network))

```

#### Nodes Attributes for Each Century In and Out Degree

```{r}

nodes_1000s<-nodes_1000s %>%
    mutate(indegree=sna::degree(wars_in_1000s_network, cmode="indegree"),
          outdegree=sna::degree(wars_in_1000s_network, cmode="outdegree"))

nodes_1100s<-nodes_1100s %>%
    mutate(indegree=sna::degree(wars_in_1100s_network, cmode="indegree"),
          outdegree=sna::degree(wars_in_1100s_network, cmode="outdegree"))

nodes_1200s<-nodes_1200s %>%
    mutate(indegree=sna::degree(wars_in_1200s_network, cmode="indegree"),
          outdegree=sna::degree(wars_in_1200s_network, cmode="outdegree"))

nodes_1300s<-nodes_1300s %>%
    mutate(indegree=sna::degree(wars_in_1300s_network, cmode="indegree"),
          outdegree=sna::degree(wars_in_1300s_network, cmode="outdegree"))

nodes_1400s<-nodes_1400s %>%
    mutate(indegree=sna::degree(wars_in_1400s_network, cmode="indegree"),
          outdegree=sna::degree(wars_in_1400s_network, cmode="outdegree"))

nodes_1500s<-nodes_1500s %>%
    mutate(indegree=sna::degree(wars_in_1500s_network, cmode="indegree"),
          outdegree=sna::degree(wars_in_1500s_network, cmode="outdegree"))

nodes_1600s<-nodes_1600s %>%
    mutate(indegree=sna::degree(wars_in_1600s_network, cmode="indegree"),
          outdegree=sna::degree(wars_in_1600s_network, cmode="outdegree"))

nodes_1700s<-nodes_1700s %>%
    mutate(indegree=sna::degree(wars_in_1700s_network, cmode="indegree"),
          outdegree=sna::degree(wars_in_1700s_network, cmode="outdegree"))

nodes_1800s<-nodes_1800s %>%
    mutate(indegree=sna::degree(wars_in_1800s_network, cmode="indegree"),
          outdegree=sna::degree(wars_in_1800s_network, cmode="outdegree"))

```

#### Eigenvector Centrality

```{r}

nodes_1000s$eigen<-evcent(wars_in_1000s_network)

nodes_1100s$eigen<-evcent(wars_in_1100s_network)

nodes_1200s$eigen<-evcent(wars_in_1200s_network)

nodes_1300s$eigen<-evcent(wars_in_1300s_network)

nodes_1400s$eigen<-evcent(wars_in_1400s_network)

nodes_1500s$eigen<-evcent(wars_in_1500s_network)

nodes_1600s$eigen<-evcent(wars_in_1600s_network)

nodes_1700s$eigen<-evcent(wars_in_1700s_network)

nodes_1800s$eigen<-evcent(wars_in_1800s_network)

```

#### Bonpow Centrality

```{r}

#nodes_1000s$bonpow <- bonpow(wars_in_1000s_network) 

#nodes_1100s$bonpow<-bonpow(wars_in_1100s_network)

#nodes_1200s$bonpow<-bonpow(wars_in_1200s_network)

#nodes_1300s$bonpow<-bonpow(wars_in_1300s_network)

#nodes_1400s$bonpow<-bonpow(wars_in_1400s_network)

#nodes_1500s$bonpow<-bonpow(wars_in_1500s_network)

#nodes_1600s$bonpow<-bonpow(wars_in_1600s_network)

#nodes_1700s$bonpow<-bonpow(wars_in_1700s_network)

#nodes_1800s$bonpow<-bonpow(wars_in_1800s_network)

```

#### Recieved and Distributed Centrality

##### 1000s rc, dc, Inital Nodes Calculations

```{r}

mat1000s<-as.matrix(as_adjacency_matrix(wars_in_1000s.ig))

mat1000s_sq<-t(mat1000s) %*% mat1000s

nodes_1000s$rc<-diag(mat1000s_sq)/rowSums(mat1000s_sq)

nodes_1000s$rc<-ifelse(is.nan(nodes_1000s$rc),0,nodes_1000s$rc)

nodes_1000s$eigen.rc<-nodes_1000s$eigen*nodes_1000s$rc

#Calculate the proportion of derived centrality.
nodes_1000s$dc<-1-diag(mat1000s_sq)/rowSums(mat1000s_sq)
#replace missing values with 0
nodes_1000s$dc<-ifelse(is.nan(nodes_1000s$dc),1,nodes_1000s$dc)
#Calculate received eigenvalue centrality
nodes_1000s$eigen.dc<-nodes_1000s$eigen*nodes_1000s$dc

```

##### 1100s rc, dc, Inital Nodes Calculations

```{r}

mat1100s<-as.matrix(as_adjacency_matrix(wars_in_1100s.ig))

mat1100s_sq<-t(mat1100s) %*% mat1100s

nodes_1100s$rc<-diag(mat1100s_sq)/rowSums(mat1100s_sq)

nodes_1100s$rc<-ifelse(is.nan(nodes_1100s$rc),0,nodes_1100s$rc)

nodes_1100s$eigen.rc<-nodes_1100s$eigen*nodes_1100s$rc

#Calculate the proportion of derived centrality.
nodes_1100s$dc<-1-diag(mat1100s_sq)/rowSums(mat1100s_sq)
#replace missing values with 0
nodes_1100s$dc<-ifelse(is.nan(nodes_1100s$dc),1,nodes_1100s$dc)
#Calculate received eigenvalue centrality
nodes_1100s$eigen.dc<-nodes_1100s$eigen*nodes_1100s$dc

```

#### 1200s rc, dc, Inital Nodes Calculations

```{r}

mat1200s<-as.matrix(as_adjacency_matrix(wars_in_1200s.ig))

mat1200s_sq<-t(mat1200s) %*% mat1200s

nodes_1200s$rc<-diag(mat1200s_sq)/rowSums(mat1200s_sq)

nodes_1200s$rc<-ifelse(is.nan(nodes_1200s$rc),0,nodes_1200s$rc)

nodes_1200s$eigen.rc<-nodes_1200s$eigen*nodes_1200s$rc

#Calculate the proportion of derived centrality.
nodes_1200s$dc<-1-diag(mat1200s_sq)/rowSums(mat1200s_sq)
#replace missing values with 0
nodes_1200s$dc<-ifelse(is.nan(nodes_1200s$dc),1,nodes_1200s$dc)
#Calculate received eigenvalue centrality
nodes_1200s$eigen.dc<-nodes_1200s$eigen*nodes_1200s$dc

```

#### 1300s rc, dc, Inital Nodes Calculations

```{r}

mat1300s<-as.matrix(as_adjacency_matrix(wars_in_1300s.ig))

mat1300s_sq<-t(mat1300s) %*% mat1300s

nodes_1300s$rc<-diag(mat1300s_sq)/rowSums(mat1300s_sq)

nodes_1300s$rc<-ifelse(is.nan(nodes_1300s$rc),0,nodes_1300s$rc)

nodes_1300s$eigen.rc<-nodes_1300s$eigen*nodes_1300s$rc

#Calculate the proportion of derived centrality.
nodes_1300s$dc<-1-diag(mat1300s_sq)/rowSums(mat1300s_sq)
#replace missing values with 0
nodes_1300s$dc<-ifelse(is.nan(nodes_1300s$dc),1,nodes_1300s$dc)
#Calculate received eigenvalue centrality
nodes_1300s$eigen.dc<-nodes_1300s$eigen*nodes_1300s$dc

```

#### 1400s rc, dc, Inital Nodes Calculations

```{r}

mat1400s<-as.matrix(as_adjacency_matrix(wars_in_1400s.ig))

mat1400s_sq<-t(mat1400s) %*% mat1400s

nodes_1400s$rc<-diag(mat1400s_sq)/rowSums(mat1400s_sq)

nodes_1400s$rc<-ifelse(is.nan(nodes_1400s$rc),0,nodes_1400s$rc)

nodes_1400s$eigen.rc<-nodes_1400s$eigen*nodes_1400s$rc

#Calculate the proportion of derived centrality.
nodes_1400s$dc<-1-diag(mat1400s_sq)/rowSums(mat1400s_sq)
#replace missing values with 0
nodes_1400s$dc<-ifelse(is.nan(nodes_1400s$dc),1,nodes_1400s$dc)
#Calculate received eigenvalue centrality
nodes_1400s$eigen.dc<-nodes_1400s$eigen*nodes_1400s$dc

```

#### 1500s rc, dc, Inital Nodes Calculations

```{r}

mat1500s<-as.matrix(as_adjacency_matrix(wars_in_1500s.ig))

mat1500s_sq<-t(mat1500s) %*% mat1500s

nodes_1500s$rc<-diag(mat1500s_sq)/rowSums(mat1500s_sq)

nodes_1500s$rc<-ifelse(is.nan(nodes_1500s$rc),0,nodes_1500s$rc)

nodes_1500s$eigen.rc<-nodes_1500s$eigen*nodes_1500s$rc

#Calculate the proportion of derived centrality.
nodes_1500s$dc<-1-diag(mat1500s_sq)/rowSums(mat1500s_sq)
#replace missing values with 0
nodes_1500s$dc<-ifelse(is.nan(nodes_1500s$dc),1,nodes_1500s$dc)
#Calculate received eigenvalue centrality
nodes_1500s$eigen.dc<-nodes_1500s$eigen*nodes_1500s$dc

```

#### 1600s rc, dc, Inital Nodes Calculations

```{r}

mat1600s<-as.matrix(as_adjacency_matrix(wars_in_1600s.ig))

mat1600s_sq<-t(mat1600s) %*% mat1600s

nodes_1600s$rc<-diag(mat1600s_sq)/rowSums(mat1600s_sq)

nodes_1600s$rc<-ifelse(is.nan(nodes_1600s$rc),0,nodes_1600s$rc)

nodes_1600s$eigen.rc<-nodes_1600s$eigen*nodes_1600s$rc

#Calculate the proportion of derived centrality.
nodes_1600s$dc<-1-diag(mat1600s_sq)/rowSums(mat1600s_sq)
#replace missing values with 0
nodes_1600s$dc<-ifelse(is.nan(nodes_1600s$dc),1,nodes_1600s$dc)
#Calculate received eigenvalue centrality
nodes_1600s$eigen.dc<-nodes_1600s$eigen*nodes_1600s$dc

```

#### 1700s rc, dc, Inital Nodes Calculations

```{r}

mat1700s<-as.matrix(as_adjacency_matrix(wars_in_1700s.ig))

mat1700s_sq<-t(mat1700s) %*% mat1700s

nodes_1700s$rc<-diag(mat1700s_sq)/rowSums(mat1700s_sq)

nodes_1700s$rc<-ifelse(is.nan(nodes_1700s$rc),0,nodes_1700s$rc)

nodes_1700s$eigen.rc<-nodes_1700s$eigen*nodes_1700s$rc

#Calculate the proportion of derived centrality.
nodes_1700s$dc<-1-diag(mat1700s_sq)/rowSums(mat1700s_sq)
#replace missing values with 0
nodes_1700s$dc<-ifelse(is.nan(nodes_1700s$dc),1,nodes_1700s$dc)
#Calculate received eigenvalue centrality
nodes_1700s$eigen.dc<-nodes_1700s$eigen*nodes_1700s$dc

```

#### 1800s rc, dc, Inital Nodes Calculations

```{r}

mat1800s<-as.matrix(as_adjacency_matrix(wars_in_1800s.ig))

mat1800s_sq<-t(mat1800s) %*% mat1800s

nodes_1800s$rc<-diag(mat1800s_sq)/rowSums(mat1800s_sq)

nodes_1800s$rc<-ifelse(is.nan(nodes_1800s$rc),0,nodes_1800s$rc)

nodes_1800s$eigen.rc<-nodes_1800s$eigen*nodes_1800s$rc

#Calculate the proportion of derived centrality.
nodes_1800s$dc<-1-diag(mat1800s_sq)/rowSums(mat1800s_sq)
#replace missing values with 0
nodes_1800s$dc<-ifelse(is.nan(nodes_1800s$dc),1,nodes_1800s$dc)
#Calculate received eigenvalue centrality
nodes_1800s$eigen.dc<-nodes_1800s$eigen*nodes_1800s$dc

```

### Brokerage and Applications in an Unusual Dataset.

```{r}

temporary_1800s_brokerage <-data.frame(brokerage(wars_in_1800s_network, cl = nodes_1800s$degree)$z.nli)

nodes_1800s <- nodes_1800s %>%
  mutate(broker.tot = temporary_1800s_brokerage$t,
         
         broker.coord = temporary_1800s_brokerage$w_I,
         
         broker.itin = temporary_1800s_brokerage$w_O,
         
         broker.rep = temporary_1800s_brokerage$b_IO,
         
         broker.gate = temporary_1800s_brokerage$b_OI,
         
         broker.lia = temporary_1800s_brokerage$b_O)


```


#### Gould Fernandez Brokerage

### Nodes Distributions 

##### Histogram 1000s Nodes Distribution

```{r}
nodes_1000s[-1] %>%
   gather() %>% 
  ggplot(aes(value)) +
    geom_histogram() +
    facet_wrap(~key, scales = "free")

```

##### Histogram 1100s Nodes Distribution

```{r}

nodes_1100s[-1] %>%
  gather() %>% 
  ggplot(aes(value)) +
    geom_histogram() +
    facet_wrap(~key, scales = "free")

```

##### Histogram 1200s Nodes Distribution

```{r}

nodes_1200s[-1] %>% 
  gather() %>% 
  ggplot(aes(value)) +
    geom_histogram() +
    facet_wrap(~key, scales = "free")

```

##### Histogram 1300s Nodes Distribution

```{r}

nodes_1300s[-1] %>%
  gather() %>% 
  ggplot(aes(value)) +
    geom_histogram() +
    facet_wrap(~key, scales = "free")

```

##### Histogram 1800s Nodes Distribution

```{r}

nodes_1400s[-1] %>% 
  gather() %>% 
  ggplot(aes(value)) +
    geom_histogram() +
    facet_wrap(~key, scales = "free")

```

##### Histogram 1500s Nodes Distribution

```{r}

nodes_1500s[-1]%>% 
  gather() %>% 
  ggplot(aes(value)) +
    geom_histogram() +
    facet_wrap(~key, scales = "free")

```

##### Histogram 1600s Nodes Distribution

```{r}

nodes_1600s[-1] %>% 
  gather() %>% 
  ggplot(aes(value)) +
    geom_histogram() +
    facet_wrap(~key, scales = "free")

```

##### Histogram 1700s Nodes Distribution

```{r}

nodes_1700s[-1] %>% 
  gather() %>% 
  ggplot(aes(value)) +
    geom_histogram() +
    facet_wrap(~key, scales = "free")

```

##### Histogram 1800s Nodes Distribution

```{r}

nodes_1800s[-1] %>% 
  gather() %>% 
  ggplot(aes(value)) +
    geom_histogram() +
    facet_wrap(~key, scales = "free")

```


```{r}

library(corrr)

```


```{r}

nodes_1800s_correlations <- nodes_1800s %>% 
  dplyr::select(degree,indegree,outdegree,eigen,eigen.rc,eigen.dc) %>%
 correlate() %>%
  rearrange()
rplot(nodes_1800s_correlations)

```

#### Centrality and Brokerage

```{r}

nodes_1000s$close <- sna::closeness(wars_in_1000s_network, cmode="suminvdir")

nodes_1100s$close <- sna::closeness(wars_in_1100s_network, cmode="suminvdir")

nodes_1200s$close <- sna::closeness(wars_in_1200s_network, cmode="suminvdir")

nodes_1300s$close <- sna::closeness(wars_in_1300s_network, cmode="suminvdir")

nodes_1400s$close <- sna::closeness(wars_in_1400s_network, cmode="suminvdir")

nodes_1500s$close <- sna::closeness(wars_in_1500s_network, cmode="suminvdir")

nodes_1600s$close <- sna::closeness(wars_in_1600s_network, cmode="suminvdir")

nodes_1700s$close <- sna::closeness(wars_in_1700s_network, cmode="suminvdir")

nodes_1800s$close <- sna::closeness(wars_in_1800s_network, cmode="suminvdir")

```


```{r}

betweenness_centralization_digraph <- c(
  
centralization(wars_in_1000s_network,FUN="betweenness",mode="digraph"),

centralization(wars_in_1100s_network,FUN="betweenness",mode="digraph"),

centralization(wars_in_1200s_network,FUN="betweenness",mode="digraph"),

centralization(wars_in_1300s_network,FUN="betweenness",mode="digraph"),

centralization(wars_in_1400s_network,FUN="betweenness",mode="digraph"),

centralization(wars_in_1500s_network,FUN="betweenness",mode="digraph"),

centralization(wars_in_1600s_network,FUN="betweenness",mode="digraph"),

centralization(wars_in_1700s_network,FUN="betweenness",mode="digraph"),

centralization(wars_in_1800s_network,FUN="betweenness",mode="digraph")

)

network_atts_for_plot$betweenness_centralization_digraph <- 
  betweenness_centralization_digraph

```


#### Closeness Centrality Over Time Graphed

```{r}


ggplot(network_atts_for_plot) +
  
geom_point(
  
  color="chocolate", 
  
  aes(
    
    x=century_for_graphing, 
                                
      y=betweenness_centralization_digraph

    )
)


```
##### Centralization

```{r}
nodes_1000s$between<-
  
sna::betweenness(wars_in_1000s_network, gmode="graph")

nodes_1100s$between<-

sna::betweenness(wars_in_1100s_network, gmode="graph")

nodes_1200s$between<-

sna::betweenness(wars_in_1200s_network, gmode="graph")

nodes_1300s$between<-

sna::betweenness(wars_in_1300s_network, gmode="graph")

nodes_1400s$between<-

sna::betweenness(wars_in_1400s_network, gmode="graph")

nodes_1500s$between<-

sna::betweenness(wars_in_1500s_network, gmode="graph")

nodes_1600s$between<-

sna::betweenness(wars_in_1600s_network, gmode="graph")

nodes_1700s$between<-

sna::betweenness(wars_in_1700s_network, gmode="graph")

nodes_1800s$between<-

sna::betweenness(wars_in_1800s_network, gmode="graph")

```

### Network Constraint 

#### Find the nodes with highest and lowest constraint

```{r}

nodes_1000s$constraint <- constraint(wars_in_1000s.ig)

nodes_1100s$constraint <- constraint(wars_in_1100s.ig)

nodes_1200s$constraint <- constraint(wars_in_1200s.ig)

nodes_1300s$constraint <- constraint(wars_in_1300s.ig)

nodes_1400s$constraint <- constraint(wars_in_1400s.ig)

nodes_1500s$constraint <- constraint(wars_in_1500s.ig)

nodes_1600s$constraint <- constraint(wars_in_1600s.ig)

nodes_1700s$constraint <- constraint(wars_in_1700s.ig)

nodes_1800s$constraint <- constraint(wars_in_1800s.ig)

```

```{r}

correlations_1800s_nodes <- nodes_1800s %>% 
  dplyr::select(degree,indegree,outdegree,eigen,close, between, broker.tot, broker.coord,broker.itin, broker.rep, broker.gate, broker.lia, constraint)%>%
  correlate() %>%
  rearrange()

rplot(correlations_1800s_nodes)

```

### Structural equivalance

#### S.E. 1000s Sedist

```{r}

wars_in_1000s.ig_no_mult <- igraph::simplify(
  wars_in_1000s.ig,
  remove.multiple = TRUE
)

wars_in_1000s_network_no_mult <- asNetwork(wars_in_1000s.ig_no_mult)

wars_in_1000s.se <- equiv.clust(wars_in_1000s_network_no_mult, equiv.fun="sedist", method="hamming", mode="digraph")

plot(wars_in_1000s.se, labels=wars_in_1000s.se$glabels,  cex=0.3)

```

#### S.E. 1000s Average

```{r}

#with average cluster.method
wars_in_1000s_network_no_mult.avg.se<-equiv.clust(wars_in_1000s_network_no_mult, equiv.fun="sedist", cluster.method="average", method="hamming", mode="digraph")

#plot:
plot(wars_in_1000s_network_no_mult.avg.se,labels=wars_in_1000s.se$glabels, cex=0.3)

```

#### S.E. 1000s Single

```{r}
#with single cluster.method
wars_in_1000s_network_no_mult.sing.se<-equiv.clust(
  wars_in_1000s_network_no_mult, equiv.fun="sedist", 
  cluster.method="single", method="hamming", mode="digraph")

#plot:
plot(wars_in_1000s_network_no_mult.sing.se, labels=wars_in_1000s.se$glabels, cex=0.3 )

```

#### S.E. 1000s Ward-D

```{r}

#with ward.D cluster.method
wars_in_1000s_network_no_mult.wrd.se<-equiv.clust(wars_in_1000s_network_no_mult, equiv.fun="sedist", cluster.method="ward.D", method="hamming",mode="digraph")

wars_in_1000s_network_no_mult
plot(wars_in_1000s_network_no_mult.wrd.se,labels=
       wars_in_1000s.se$glabels, cex=0.3)

```

#### Adding Degree Weight

```{r}

nodes_1800s<-nodes_1800s %>% data.frame(
  
degree.wt=strength(wars_in_1800s.ig)

)

```

### Community Detection

#### Fast and Greedy Community Detection 1000s

Fast Greedy does not support directed graphs so I have to convert my wars object to an undirected graph making this likely a weaker method for community detection.

```{r}

# The Giant Component below is loaded here but factors into community detection
giant.component <- function(graph) {
  cl <- clusters(graph)
  induced.subgraph(graph, which(cl$membership == which.max(cl$csize)))
}
```

```{r}

wars_in_1000s_undirected.ig <- igraph::as.undirected(
  wars_in_1000s.ig,
  mode = c("collapse", "each", "mutual"),
)

wars_in_1000s_comm.fg<-cluster_fast_greedy(wars_in_1000s_undirected.ig)

names(wars_in_1000s_comm.fg)

```

#### Fast and Greedy Blockmodel 1000s

```{r}


wars_in_1000s_undirected_network <- asNetwork(wars_in_1000s_undirected.ig)

wars_in_1000s_undirected_network_blockmodel <- 
  blockmodel(wars_in_1000s_undirected_network, wars_in_1000s_comm.fg$membership)


```

#### Node Attributes Fast Greedy Membership 1000s

```{r}

#add community membership as a node attribute
nodes_1000s$comm.fg <- wars_in_1000s_comm.fg$membership
#summarize node statistics by community

```

#### Walktrap Community Detection 1000s

```{r}
#Run clustering algorithm: walktrap
wars_in_1000s_comm.wt <- walktrap.community(wars_in_1000s.ig)
#add community membership as a vertex attribute
nodes_1000s$comm.wt <- wars_in_1000s_comm.wt$membership
#summarize node statistics by community
```

#### Leading Label Propagation Community Detection 1000s

```{r}

wars_in_1000s_comm.lab <- label.propagation.community(wars_in_1000s.ig)

nodes_1000s$comm.lab<-wars_in_1000s_comm.lab$membership

```

#### Edge Betweenness Community Detection 1000s

```{r}

wars_in_1000s_comm.edge<-label.propagation.community(wars_in_1000s.ig)

nodes_1000s$comm.edge<-wars_in_1000s_comm.edge$membership

```

#### Eigenvector Community Detection 1000s

```{r}

wars_in_1000s_comm.eigen <- leading.eigenvector.community(wars_in_1000s_undirected.ig)

nodes_1000s$comm.eigen<-wars_in_1000s_comm.eigen$membership

```

#### Fast and Greedy Community Detection 1100s

Fast Greedy does not support directed graphs so I have to convert my wars object to an undirected graph making this likely a weaker method for community detection.

```{r}

wars_in_1100s_undirected.ig <- igraph::as.undirected(
  wars_in_1100s.ig,
  mode = c("collapse", "each", "mutual"),
)

wars_in_1100s_comm.fg<-cluster_fast_greedy(wars_in_1100s_undirected.ig)

names(wars_in_1100s_comm.fg)

```

#### Fast and Greedy Blockmodel 1100s

```{r}


wars_in_1100s_undirected_network <- asNetwork(wars_in_1100s_undirected.ig)

wars_in_1100s_undirected_network_blockmodel <- blockmodel(wars_in_1100s_undirected_network,wars_in_1100s_comm.fg$membership)


```

#### Node Attributes Fast Greedy Membership 1100s

```{r}

#add community membership as a node attribute
nodes_1100s$comm.fg<-wars_in_1100s_comm.fg$membership
#summarize node statistics by community

```

#### Walktrap Community Detection 1100s

```{r}
#Run clustering algorithm: walktrap
wars_in_1100s_comm.wt<-walktrap.community(wars_in_1100s.ig)
#add community membership as a vertex attribute
nodes_1100s$comm.wt <- wars_in_1100s_comm.wt$membership
#summarize node statistics by community
```

#### Leading Label Propagation Community Detection 1100s

```{r}

wars_in_1100s_comm.lab<-label.propagation.community(wars_in_1100s.ig)

nodes_1100s$comm.lab<-wars_in_1100s_comm.lab$membership

```

#### Edge Betweenness Community Detection 1100s

```{r}

wars_in_1100s_comm.edge<-label.propagation.community(wars_in_1100s.ig)

nodes_1100s$comm.edge<-wars_in_1100s_comm.edge$membership

```

#### Eigenvector Community Detection 1100s

```{r}

wars_in_1100s_comm.eigen <- leading.eigenvector.community(wars_in_1100s_undirected.ig)

nodes_1100s$comm.eigen<-wars_in_1100s_comm.eigen$membership

```

#### Fast and Greedy Community Detection 1200s

Fast Greedy does not support directed graphs so I have to convert my wars object to an undirected graph making this likely a weaker method for community detection.

```{r}

wars_in_1200s_undirected.ig <- igraph::as.undirected(
  wars_in_1200s.ig,
  mode = c("collapse", "each", "mutual"),
)

wars_in_1200s_comm.fg<-cluster_fast_greedy(wars_in_1200s_undirected.ig)

names(wars_in_1200s_comm.fg)

```

#### Fast and Greedy Blockmodel 1200s

```{r}


wars_in_1200s_undirected_network <- asNetwork(wars_in_1200s_undirected.ig)

wars_in_1200s_undirected_network_blockmodel <- blockmodel(wars_in_1200s_undirected_network,wars_in_1200s_comm.fg$membership)


```

#### Node Attributes Fast Greedy Membership 1200s

```{r}

#add community membership as a node attribute
nodes_1200s$comm.fg<-wars_in_1200s_comm.fg$membership
#summarize node statistics by community

```

#### Walktrap Community Detection 1200s

```{r}
#Run clustering algorithm: walktrap
wars_in_1200s_comm.wt<-walktrap.community(wars_in_1200s.ig)
#add community membership as a vertex attribute
nodes_1200s$comm.wt <- wars_in_1200s_comm.wt$membership
#summarize node statistics by community
```

#### Leading Label Propagation Community Detection 1200s

```{r}

wars_in_1200s_comm.lab<-label.propagation.community(wars_in_1200s.ig)

nodes_1200s$comm.lab<-wars_in_1200s_comm.lab$membership

```

#### Edge Betweenness Community Detection 1200s

```{r}

wars_in_1200s_comm.edge<-label.propagation.community(wars_in_1200s.ig)

nodes_1200s$comm.edge<-wars_in_1200s_comm.edge$membership

```

#### Eigenvector Community Detection 1200s

**Big Asterisk here!**

```{r}

wars_in_1200s_comm.eigen <- leading.eigenvector.community(wars_in_1200s.ig)

nodes_1200s$comm.eigen<-wars_in_1200s_comm.eigen$membership

```

#### Fast and Greedy Community Detection 1300s

Fast Greedy does not support directed graphs so I have to convert my wars object to an undirected graph making this likely a weaker method for community detection.

```{r}

wars_in_1300s_undirected.ig <- igraph::as.undirected(
  wars_in_1300s.ig,
  mode = c("collapse", "each", "mutual"),
)

wars_in_1300s_comm.fg<-cluster_fast_greedy(wars_in_1300s_undirected.ig)

names(wars_in_1300s_comm.fg)

```

#### Fast and Greedy Blockmodel 1300s

```{r}


wars_in_1300s_undirected_network <- asNetwork(wars_in_1300s_undirected.ig)

wars_in_1300s_undirected_network_blockmodel <- blockmodel(wars_in_1300s_undirected_network, wars_in_1300s_comm.fg$membership)


```

#### Node Attributes Fast Greedy Membership 1300s

```{r}

#add community membership as a node attribute
nodes_1300s$comm.fg<-wars_in_1300s_comm.fg$membership
#summarize node statistics by community

```

#### Walktrap Community Detection 1300s

```{r}
#Run clustering algorithm: walktrap
wars_in_1300s_comm.wt<-walktrap.community(wars_in_1300s.ig)
#add community membership as a vertex attribute
nodes_1300s$comm.wt <- wars_in_1300s_comm.wt$membership
#summarize node statistics by community
```

#### Leading Label Propagation Community Detection 1300s

```{r}

wars_in_1300s_comm.lab<-label.propagation.community(wars_in_1300s.ig)

nodes_1300s$comm.lab<-wars_in_1300s_comm.lab$membership

```

#### Edge Betweenness Community Detection 1300s

```{r}

wars_in_1300s_comm.edge<-label.propagation.community(wars_in_1300s.ig)

nodes_1300s$comm.edge<-wars_in_1300s_comm.edge$membership

```

#### Eigenvector Community Detection 1300s

```{r}

wars_in_1300s_comm.eigen <- leading.eigenvector.community(wars_in_1300s_undirected.ig)

nodes_1300s$comm.eigen<-wars_in_1300s_comm.eigen$membership

```

#### Fast and Greedy Community Detection 1400s

Fast Greedy does not support directed graphs so I have to convert my wars object to an undirected graph making this likely a weaker method for community detection.

```{r}

wars_in_1400s_undirected.ig <- igraph::as.undirected(
  wars_in_1400s.ig,
  mode = c("collapse", "each", "mutual"),
)

wars_in_1400s_comm.fg<-cluster_fast_greedy(wars_in_1400s_undirected.ig)

names(wars_in_1400s_comm.fg)

```

#### Fast and Greedy Blockmodel 1400s

```{r}


wars_in_1400s_undirected_network <- asNetwork(wars_in_1400s_undirected.ig)

wars_in_1400s_undirected_network_blockmodel <- blockmodel(wars_in_1400s_undirected_network, wars_in_1400s_comm.fg$membership)


```

#### Node Attributes Fast Greedy Membership 1400s

```{r}

#add community membership as a node attribute
nodes_1400s$comm.fg<-wars_in_1400s_comm.fg$membership
#summarize node statistics by community

```

#### Walktrap Community Detection 1400s

```{r}
#Run clustering algorithm: walktrap
wars_in_1400s_comm.wt<-walktrap.community(wars_in_1400s.ig)
#add community membership as a vertex attribute
nodes_1400s$comm.wt <- wars_in_1400s_comm.wt$membership
#summarize node statistics by community
```

#### Leading Label Propagation Community Detection 1400s

```{r}

wars_in_1400s_comm.lab<-label.propagation.community(wars_in_1400s.ig)

nodes_1400s$comm.lab<-wars_in_1400s_comm.lab$membership

```

#### Edge Betweenness Community Detection 1400s

```{r}

wars_in_1400s_comm.edge<-label.propagation.community(wars_in_1400s.ig)

nodes_1400s$comm.edge<-wars_in_1400s_comm.edge$membership

```

#### Eigenvector Community Detection 1400s

**Big Asterisk here!**

```{r}

wars_in_1400s_comm.eigen <- leading.eigenvector.community(wars_in_1400s.ig)

nodes_1400s$comm.eigen<-wars_in_1400s_comm.eigen$membership

```

#### Fast and Greedy Community Detection 1500s

Fast Greedy does not support directed graphs so I have to convert my wars object to an undirected graph making this likely a weaker method for community detection.

```{r}

wars_in_1500s_undirected.ig <- igraph::as.undirected(
  wars_in_1500s.ig,
  mode = c("collapse", "each", "mutual"),
)

wars_in_1500s_comm.fg<-cluster_fast_greedy(wars_in_1500s_undirected.ig)

names(wars_in_1500s_comm.fg)

```

#### Fast and Greedy Blockmodel 1500s

```{r}


wars_in_1500s_undirected_network <- asNetwork(wars_in_1500s_undirected.ig)

wars_in_1500s_undirected_network_blockmodel <- blockmodel(wars_in_1500s_undirected_network, wars_in_1500s_comm.fg$membership)


```

#### Node Attributes Fast Greedy Membership 1500s

```{r}

#add community membership as a node attribute
nodes_1500s$comm.fg<-wars_in_1500s_comm.fg$membership
#summarize node statistics by community

```

#### Walktrap Community Detection 1500s

```{r}
#Run clustering algorithm: walktrap
wars_in_1500s_comm.wt<-walktrap.community(wars_in_1500s.ig)
#add community membership as a vertex attribute
nodes_1500s$comm.wt <- wars_in_1500s_comm.wt$membership
#summarize node statistics by community
```

#### Leading Label Propagation Community Detection 1500s

```{r}

wars_in_1500s_comm.lab<-label.propagation.community(wars_in_1500s.ig)

nodes_1500s$comm.lab<-wars_in_1500s_comm.lab$membership

```

#### Edge Betweenness Community Detection 1500s

```{r}

wars_in_1500s_comm.edge<-label.propagation.community(wars_in_1500s.ig)

nodes_1500s$comm.edge<-wars_in_1500s_comm.edge$membership

```

#### Eigenvector Community Detection 1500s

**Big Asterisk here!**

```{r}

wars_in_1500s_comm.eigen <- leading.eigenvector.community(wars_in_1500s.ig)

nodes_1500s$comm.eigen<-wars_in_1500s_comm.eigen$membership

```

#### Fast and Greedy Community Detection 1600s

Fast Greedy does not support directed graphs so I have to convert my wars object to an undirected graph making this likely a weaker method for community detection.

```{r}

wars_in_1600s_undirected.ig <- igraph::as.undirected(
  wars_in_1600s.ig,
  mode = c("collapse", "each", "mutual"),
)

wars_in_1600s_comm.fg<-cluster_fast_greedy(wars_in_1600s_undirected.ig)

names(wars_in_1600s_comm.fg)

```

#### Fast and Greedy Blockmodel 1600s

```{r}


wars_in_1600s_undirected_network <- asNetwork(wars_in_1600s_undirected.ig)

wars_in_1600s_undirected_network_blockmodel <- blockmodel(wars_in_1600s_undirected_network, wars_in_1600s_comm.fg$membership)


```

#### Node Attributes Fast Greedy Membership 1600s

```{r}

#add community membership as a node attribute
nodes_1600s$comm.fg<-wars_in_1600s_comm.fg$membership
#summarize node statistics by community

```

#### Walktrap Community Detection 1600s

```{r}
#Run clustering algorithm: walktrap
wars_in_1600s_comm.wt<-walktrap.community(wars_in_1600s.ig)
#add community membership as a vertex attribute
nodes_1600s$comm.wt <- wars_in_1600s_comm.wt$membership
#summarize node statistics by community
```

#### Leading Label Propagation Community Detection 1600s

```{r}

wars_in_1600s_comm.lab<-label.propagation.community(wars_in_1600s.ig)

nodes_1600s$comm.lab<-wars_in_1600s_comm.lab$membership

```

#### Edge Betweenness Community Detection 1600s

```{r}

wars_in_1600s_comm.edge<-label.propagation.community(wars_in_1600s.ig)

nodes_1600s$comm.edge<-wars_in_1600s_comm.edge$membership

```

#### Eigenvector Community Detection 1600s

```{r}

wars_in_1600s_comm.eigen <- leading.eigenvector.community(wars_in_1600s_undirected.ig)

nodes_1600s$comm.eigen<-wars_in_1600s_comm.eigen$membership

```

#### Fast and Greedy Community Detection 1700s

Fast Greedy does not support directed graphs so I have to convert my wars object to an undirected graph making this likely a weaker method for community detection.

```{r}

wars_in_1700s_undirected.ig <- igraph::as.undirected(
  wars_in_1700s.ig,
  mode = c("collapse", "each", "mutual"),
)

wars_in_1700s_comm.fg<-cluster_fast_greedy(wars_in_1700s_undirected.ig)

names(wars_in_1700s_comm.fg)

```

#### Fast and Greedy Blockmodel 1700s

```{r}


wars_in_1700s_undirected_network <- asNetwork(wars_in_1700s_undirected.ig)

wars_in_1700s_undirected_network_blockmodel <- blockmodel(wars_in_1700s_undirected_network, wars_in_1700s_comm.fg$membership)


```

#### Node Attributes Fast Greedy Membership 1700s

```{r}

#add community membership as a node attribute
nodes_1700s$comm.fg<-wars_in_1700s_comm.fg$membership
#summarize node statistics by community

```

#### Walktrap Community Detection 1700s

```{r}
#Run clustering algorithm: walktrap
wars_in_1700s_comm.wt<-walktrap.community(wars_in_1700s.ig)
#add community membership as a vertex attribute
nodes_1700s$comm.wt <- wars_in_1700s_comm.wt$membership
#summarize node statistics by community
```

#### Leading Label Propagation Community Detection 1700s

```{r}

wars_in_1700s_comm.lab<-label.propagation.community(wars_in_1700s.ig)

nodes_1700s$comm.lab<-wars_in_1700s_comm.lab$membership

```

#### Edge Betweenness Community Detection 1700s

```{r}

wars_in_1700s_comm.edge<-label.propagation.community(wars_in_1700s.ig)

nodes_1700s$comm.edge<-wars_in_1700s_comm.edge$membership

```

#### Eigenvector Community Detection 1700s

```{r}

wars_in_1700s_comm.eigen <- leading.eigenvector.community(wars_in_1700s_undirected.ig)

nodes_1700s$comm.eigen<-wars_in_1700s_comm.eigen$membership

```

#### Fast and Greedy Community Detection 1800s

Fast Greedy does not support directed graphs so I have to convert my wars object to an undirected graph making this likely a weaker method for community detection.

```{r}

wars_in_1800s_undirected.ig <- igraph::as.undirected(
  wars_in_1800s.ig,
  mode = c("collapse", "each", "mutual"),
)

wars_in_1800s_comm.fg<-cluster_fast_greedy(wars_in_1800s_undirected.ig)

names(wars_in_1800s_comm.fg)

```

#### Fast and Greedy Blockmodel 1800s

```{r}


wars_in_1800s_undirected_network <- asNetwork(wars_in_1800s_undirected.ig)

wars_in_1800s_undirected_network_blockmodel <- blockmodel(wars_in_1800s_undirected_network, wars_in_1800s_comm.fg$membership)


```

#### Node Attributes Fast Greedy Membership 1800s

```{r}

#add community membership as a node attribute
nodes_1800s$comm.fg<-wars_in_1800s_comm.fg$membership
#summarize node statistics by community

```

#### Walktrap Community Detection 1800s

```{r}
#Run clustering algorithm: walktrap
wars_in_1800s_comm.wt<-walktrap.community(wars_in_1800s.ig)
#add community membership as a vertex attribute
nodes_1800s$comm.wt <- wars_in_1800s_comm.wt$membership
#summarize node statistics by community
```

#### Leading Label Propagation Community Detection 1800s

```{r}

wars_in_1800s_comm.lab<-label.propagation.community(wars_in_1800s.ig)

nodes_1800s$comm.lab<-wars_in_1800s_comm.lab$membership

```

#### Edge Betweenness Community Detection 1800s

```{r}

wars_in_1800s_comm.edge<-label.propagation.community(wars_in_1800s.ig)

nodes_1800s$comm.edge<-wars_in_1800s_comm.edge$membership

```

#### Eigenvector Community Detection 1800s

```{r}

wars_in_1800s_comm.eigen <- leading.eigenvector.community(wars_in_1800s_undirected.ig)

nodes_1800s$comm.eigen<-wars_in_1800s_comm.eigen$membership

```

#### Comparing Modualrity Scores 

```{r}
compare.algs<-function(alg.a,alg.b,compare.meth=c("vi", "nmi", "split.join", "rand", "adjusted.rand")){
  #create list of community objects and methods
  comm.compare<-expand.grid(alg.a=alg.a, alg.b=alg.b, meth=compare.meth, result=NA, stringsAsFactors = FALSE)
  #compare community partitions using a loop
for(i in 1:nrow(comm.compare)){
  comm1<-get(comm.compare$alg.a[i])
  comm2<-get(comm.compare$alg.b[i])
  method<-comm.compare$meth[i]
  comm.compare$result[i]<-compare(comm1, comm2, method)
}
  return(comm.compare)
}
```

#### Comparing 1000s Modualrity Scores 

```{r}

compare.algs(alg.a=c("wars_in_1000s_comm.fg","wars_in_1000s_comm.wt", "wars_in_1000s_comm.lab", "wars_in_1000s_comm.edge"), alg.b="wars_in_1000s_comm.eigen")

```
#### Comparing 1100s Modualrity Scores 

```{r}

compare.algs(alg.a=c("wars_in_1100s_comm.fg","wars_in_1100s_comm.wt", "wars_in_1100s_comm.lab", "wars_in_1100s_comm.edge"), alg.b="wars_in_1100s_comm.eigen")

```

#### Comparing 1200s Modualrity Scores 

```{r}

compare.algs(alg.a=c("wars_in_1200s_comm.fg","wars_in_1200s_comm.wt", "wars_in_1200s_comm.lab", "wars_in_1200s_comm.edge"), alg.b="wars_in_1200s_comm.eigen")

```

#### Comparing 1300s Modualrity Scores 

```{r}

compare.algs(alg.a=c("wars_in_1300s_comm.fg","wars_in_1300s_comm.wt", "wars_in_1300s_comm.lab", "wars_in_1300s_comm.edge"), alg.b="wars_in_1300s_comm.eigen")

```

#### Comparing 1400s Modualrity Scores 

```{r}

compare.algs(alg.a=c("wars_in_1400s_comm.fg","wars_in_1400s_comm.wt", "wars_in_1400s_comm.lab", "wars_in_1400s_comm.edge"), alg.b="wars_in_1400s_comm.eigen")

```

#### Comparing 1500s Modualrity Scores 

```{r}

compare.algs(alg.a=c("wars_in_1500s_comm.fg","wars_in_1500s_comm.wt", "wars_in_1500s_comm.lab", "wars_in_1500s_comm.edge"), alg.b="wars_in_1500s_comm.eigen")

```

#### Comparing 1600s Modualrity Scores 

```{r}

compare.algs(alg.a=c("wars_in_1600s_comm.fg","wars_in_1600s_comm.wt", "wars_in_1600s_comm.lab", "wars_in_1600s_comm.edge"), alg.b="wars_in_1600s_comm.eigen")

```

#### Comparing 1700s Modualrity Scores 

```{r}

compare.algs(alg.a=c("wars_in_1700s_comm.fg","wars_in_1700s_comm.wt", "wars_in_1700s_comm.lab", "wars_in_1700s_comm.edge"), alg.b="wars_in_1700s_comm.eigen")

```

#### Comparing 1800s Modualrity Scores 

```{r}

compare.algs(alg.a=c("wars_in_1800s_comm.fg","wars_in_1800s_comm.wt", "wars_in_1800s_comm.lab", "wars_in_1800s_comm.edge"), alg.b="wars_in_1800s_comm.eigen")

```

### Histograms of Important Network Attributes 

##### Histogram 1000s Degree Distribution

```{r}

hist(nodes_1000s$indegree, main="Conflict in 1000s: In-degree Distribution", 
     xlab="Co-Belligerents Defeated")

hist(nodes_1000s$outdegree, main="Conflict in 1000s: Out-degree Distribution", 
     xlab="Co-Belligerents Lost to")

```

##### Histogram 1100s Degree Distribution

```{r}

hist(nodes_1100s$indegree, main="Conflict in 1100s: In-degree Distribution", 
     xlab="Co-Belligerents Defeated")

hist(nodes_1100s$outdegree, main="Conflict in 1100s: Out-degree Distribution", 
     xlab="Co-Belligerents Lost to")

```

##### Histogram 1200s Degree Distribution

```{r}

hist(nodes_1200s$indegree, main="Conflict in 1200s: In-degree Distribution", 
     xlab="Co-Belligerents Defeated")

hist(nodes_1200s$outdegree, main="Conflict in 1200s: Out-degree Distribution", 
     xlab="Co-Belligerents Lost to")

```

##### Histogram 1300s Degree Distribution

```{r}

hist(nodes_1300s$indegree, main="Conflict in 1300s: In-degree Distribution",
     xlab="Co-Belligerents Defeated")

hist(nodes_1300s$outdegree, main="Conflict in 1300s: Out-degree Distribution", 
     xlab="Co-Belligerents Lost to")

```

##### Histogram 1400s Degree Distribution

```{r}

hist(nodes_1400s$indegree, main="Conflict in 1400s: In-degree Distribution", 
     xlab="Co-Belligerents Defeated")

hist(nodes_1400s$outdegree, main="Conflict in 1400s: Out-degree Distribution", 
     xlab="Co-Belligerents Lost to")

```

##### Histogram 1500s Degree Distribution

```{r}

hist(nodes_1500s$indegree, main="Conflict in 1500s: In-degree Distribution", 
     xlab="Co-Belligerents Defeated")

hist(nodes_1500s$outdegree, main="Conflict in 1500s: Out-degree Distribution", 
     xlab="Co-Belligerents Lost to")

```

##### Histogram 1600s Degree Distribution

```{r}

hist(nodes_1600s$indegree, main="Conflict in 1600s: In-degree Distribution", 
     xlab="Co-Belligerents Defeated")

hist(nodes_1600s$outdegree, main="Conflict in 1600s: Out-degree Distribution", 
     xlab="Co-Belligerents Lost to")

```

##### Histogram 1700s Degree Distribution

```{r}

hist(nodes_1700s$indegree, main="Conflict in 1700s: In-degree Distribution", 
     xlab="Co-Belligerents Defeated")

hist(nodes_1700s$outdegree, main="Conflict in 1700s: Out-degree Distribution", 
     xlab="Co-Belligerents Lost to")

```

##### Histogram 1800s Degree Distribution

```{r}

hist(nodes_1800s$indegree, main="Conflict in 1800s: In-degree Distribution", 
     xlab="Co-Belligerents Defeated")

hist(nodes_1800s$outdegree, main="Conflict in 1800s: Out-degree Distribution", 
     xlab="Co-Belligerents Lost to")

```

```{r}

set.seed(777)
ggplot(wars_in_1000s_network, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "grey25", alpha = 0.25, arrow = arrow(length = unit(2, "pt"), type = "closed"))+
  geom_nodes(color = "purple1", alpha = 0.3, size = 2) +
  geom_nodetext(aes( label = vertex.names ), size=1)+
    ggtitle("Networks of Wars in the 1000s") +
  theme_blank()
```

```{r}

set.seed(777)
ggplot(wars_in_1100s_network, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "grey25", alpha = 0.25, arrow = arrow(length = unit(2, "pt"), type = "closed"))+
  geom_nodes(color = "gold", alpha = 0.3, size = 2) +
  geom_nodetext(aes( label = vertex.names ), size=1)+
    ggtitle("Networks of Wars in the 1100s") +
  theme_blank()
```

```{r}

set.seed(777)
ggplot(wars_in_1200s_network, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "grey25", alpha = 0.25, arrow = arrow(length = unit(2, "pt"), type = "closed"))+
  geom_nodes(color = "forestgreen", alpha = 0.3, size = 2) +
  geom_nodetext(aes( label = vertex.names ), size=1)+
    ggtitle("Networks of Wars in the 1200s") +
  theme_blank()
```

```{r}

set.seed(777)
ggplot(wars_in_1300s_network, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "grey25", alpha = 0.25, arrow = arrow(length = unit(2, "pt"), type = "closed"))+
  geom_nodes(color = "burlywood1", alpha = 0.3, size = 2) +
  geom_nodetext(aes( label = vertex.names ), size=1)+
    ggtitle("Networks of Wars in the 1300s") +
  theme_blank()
```

```{r}

set.seed(777)
ggplot(wars_in_1400s_network, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "grey25", alpha = 0.25, arrow = arrow(length = unit(2, "pt"), type = "closed"))+
  geom_nodes(color = "aquamarine2", alpha = 0.3, size = 2) +
  geom_nodetext(aes( label = vertex.names ), size=1)+
    ggtitle("Networks of Wars in the 1400s") +
  theme_blank()
```

```{r}

set.seed(777)
ggplot(wars_in_1500s_network, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "grey25", alpha = 0.25, arrow = arrow(length = unit(2, "pt"), type = "closed"))+
  geom_nodes(color = "mediumorchid1", alpha = 0.3, size = 2) +
  geom_nodetext(aes( label = vertex.names ), size=1)+
    ggtitle("Networks of Wars in the 1500s") +
  theme_blank()
```

```{r}

set.seed(777)
ggplot(wars_in_1600s_network, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "grey25", alpha = 0.25, arrow = arrow(length = unit(2, "pt"), type = "closed"))+
  geom_nodes(color = "plum2", alpha = 0.3, size = 2) +
  geom_nodetext(aes( label = vertex.names ), size=1)+
    ggtitle("Networks of Wars in the 1600s") +
  theme_blank()
```

```{r}

set.seed(777)
ggplot(wars_in_1700s_network, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "grey25", alpha = 0.25, arrow = arrow(length = unit(2, "pt"), type = "closed"))+
  geom_nodes(color = "tomato1", alpha = 0.3, size = 2) +
  geom_nodetext(aes( label = vertex.names ), size=1)+
    ggtitle("Networks of Wars in the 1700s") +
  theme_blank()
```

```{r}

set.seed(777)
ggplot(wars_in_1800s_network, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "grey25", alpha = 0.25, arrow = arrow(length = unit(2, "pt"), type = "closed"))+
  geom_nodes(color = "steelblue1", alpha = 0.3, size = 2) +
  geom_nodetext(aes( label = vertex.names ), size=1)+
    ggtitle("Networks of Wars in the 1800s") +
  theme_blank()
```

```{r}

set.seed(238)
#Run clustering algorithm: fast_greedy
wars_in_1000s.wt<-walktrap.community(wars_in_1000s.ig)

#igraph::groups(wars_in_1000s.wt)

```

Adding more steps resulted in 19 groups for both 10 and 20 steps.

```{r}
#Run & inspect clustering algorithm: 10 steps
#igraph::groups(walktrap.community(wars_in_1000s.ig, steps=10)) 
#Run & inspect clustering algorithm: 20 steps
#igraph::groups(walktrap.community(wars_in_1000s.ig ,steps=20))
#Run & inspect clustering algorithm
```

```{r, echo=FALSE}

plot(wars_in_1000s.wt, wars_in_1000s.ig, vertex.size=0.5, edge.size=0.25, edge.arrow.size=.25,  vertex.label.dist=1.5,vertex.label.cex=0.25,asp = 0)

```
```{r}

set.seed(238)
#Run clustering algorithm: fast_greedy
wars_in_1100s.wt<-walktrap.community(wars_in_1100s.ig)

#igraph::groups(wars_in_1000s.wt)

```

```{r, echo=FALSE}

plot(wars_in_1100s.wt, wars_in_1100s.ig, vertex.size=0.5, edge.size=0.25, edge.arrow.size=.25,  vertex.label.dist=1.5,vertex.label.cex=0.25,asp = 0)

```

```{r}

set.seed(238)
#Run clustering algorithm: fast_greedy
wars_in_1200s.wt<-walktrap.community(wars_in_1200s.ig)

#igraph::groups(wars_in_1000s.wt)

```

```{r, echo=FALSE}

plot(wars_in_1200s.wt, wars_in_1200s.ig, vertex.size=0.5, edge.size=0.25, edge.arrow.size=.25,  vertex.label.dist=1.5,vertex.label.cex=0.25,asp = 0)

```

```{r}

set.seed(238)
#Run clustering algorithm: fast_greedy
wars_in_1300s.wt<-walktrap.community(wars_in_1300s.ig)

#igraph::groups(wars_in_1000s.wt)

```

```{r, echo=FALSE}

plot(wars_in_1300s.wt, wars_in_1300s.ig, vertex.size=0.5, edge.size=0.25, edge.arrow.size=.25,  vertex.label.dist=1.5,vertex.label.cex=0.25,asp = 0)

```

```{r}

set.seed(238)
#Run clustering algorithm: fast_greedy
wars_in_1400s.wt<-walktrap.community(wars_in_1400s.ig)

#igraph::groups(wars_in_1000s.wt)

```

```{r, echo=FALSE}

plot(wars_in_1400s.wt, wars_in_1400s.ig, vertex.size=0.5, edge.size=0.25, edge.arrow.size=.25,  vertex.label.dist=1.5,vertex.label.cex=0.25,asp = 0)

```

```{r}

set.seed(238)
#Run clustering algorithm: fast_greedy
wars_in_1500s.wt<-walktrap.community(wars_in_1500s.ig)

#igraph::groups(wars_in_1000s.wt)

```

```{r, echo=FALSE}

plot(wars_in_1500s.wt, wars_in_1500s.ig, vertex.size=0.5, edge.size=0.25, edge.arrow.size=.25,  vertex.label.dist=1.5,vertex.label.cex=0.25,asp = 0)

```
```{r}

set.seed(238)
#Run clustering algorithm: fast_greedy
wars_in_1600s.wt<-walktrap.community(wars_in_1600s.ig)

#igraph::groups(wars_in_1000s.wt)

```

```{r, echo=FALSE}

plot(wars_in_1600s.wt, wars_in_1600s.ig, vertex.size=0.5, edge.size=0.25, edge.arrow.size=.25,  vertex.label.dist=1.5,vertex.label.cex=0.25,asp = 0)

```

```{r}

set.seed(238)
#Run clustering algorithm: fast_greedy
wars_in_1700s.wt<-walktrap.community(wars_in_1700s.ig)

#igraph::groups(wars_in_1000s.wt)

```

```{r, echo=FALSE}

plot(wars_in_1700s.wt, wars_in_1700s.ig, vertex.size=0.5, edge.size=0.25, edge.arrow.size=.25,  vertex.label.dist=1.5,vertex.label.cex=0.25,asp = 0)

```

```{r}

set.seed(238)
#Run clustering algorithm: fast_greedy
wars_in_1800s.wt<-walktrap.community(wars_in_1800s.ig)

#igraph::groups(wars_in_1000s.wt)

```

```{r, echo=FALSE}

plot(wars_in_1800s.wt, wars_in_1800s.ig, vertex.size=0.5, edge.size=0.25, edge.arrow.size=.25,  vertex.label.dist=1.5,vertex.label.cex=0.25,asp = 0)

```

```{r}
library(ape)
library(ggdendro)
library(sets)
library(ggplot2)
library(dendextend)
library(RColorBrewer)

```

```{r}

#pdf("h_clust_plot.pdf") 

dd <- dist(scale(wars_1000s), method = "euclidean")

hc <- hclust(dd, method = "complete")

dend <- hc %>% as.dendrogram() %>%
   set("branches_k_color", k=39) %>% 
  set("branches_lwd", 1)  %>%     
  set("labels_cex", c(.3,0.3)) 

# plot the dend in usual "base" plotting engine:
plot(dend,col="white", type="rectangle", frame.plot = FALSE, center	=FALSE, xlab="Belligerents", ylab="Distance (Euclidean)", main="Complete, Euclidean Distance Dendrogram" , cex=0.5,   horiz = FALSE)



```

```{r}

#pdf("h_clust_plot.pdf") 

dd <- dist(scale(wars_1100s), method = "euclidean")

hc <- hclust(dd, method = "complete")

dend <- hc %>% as.dendrogram() %>%
   set("branches_k_color", k=39) %>% 
  set("branches_lwd", 1)  %>%     
  set("labels_cex", c(.2,0.2)) 

# plot the dend in usual "base" plotting engine:
plot(dend,col="white", type="rectangle", frame.plot = FALSE, center	=FALSE, xlab="Belligerents", ylab="Distance (Euclidean)", main="Complete, Euclidean Distance Dendrogram" , cex=0.5,   horiz = FALSE)



```

```{r}

#pdf("h_clust_plot.pdf") 

dd <- dist(scale(wars_1200s), method = "euclidean")

hc <- hclust(dd, method = "complete")

dend <- hc %>% as.dendrogram() %>%
   set("branches_k_color", k=39) %>% 
  set("branches_lwd", 1)  %>%     
  set("labels_cex", c(.2,0.2)) 

# plot the dend in usual "base" plotting engine:
plot(dend,col="white", type="rectangle", frame.plot = FALSE, center	=FALSE, xlab="Belligerents", ylab="Distance (Euclidean)", main="Complete, Euclidean Distance Dendrogram" , cex=0.5,   horiz = FALSE)



```

```{r}

#pdf("h_clust_plot.pdf") 

dd <- dist(scale(wars_1300s), method = "euclidean")

hc <- hclust(dd, method = "complete")

dend <- hc %>% as.dendrogram() %>%
   set("branches_k_color", k=39) %>% 
  set("branches_lwd", 1)  %>%     
  set("labels_cex", c(.2,0.2)) 

# plot the dend in usual "base" plotting engine:
plot(dend,col="white", type="rectangle", frame.plot = FALSE, center	=FALSE, xlab="Belligerents", ylab="Distance (Euclidean)", main="Complete, Euclidean Distance Dendrogram" , cex=0.5,   horiz = FALSE)

```

```{r}

#pdf("h_clust_plot.pdf") 

dd <- dist(scale(wars_1400s), method = "euclidean")

hc <- hclust(dd, method = "complete")

dend <- hc %>% as.dendrogram() %>%
   set("branches_k_color", k=39) %>% 
  set("branches_lwd", 1)  %>%     
  set("labels_cex", c(.15,0.15)) 

# plot the dend in usual "base" plotting engine:
plot(dend,col="white", type="rectangle", frame.plot = FALSE, center	=FALSE, xlab="Belligerents", ylab="Distance (Euclidean)", main="Complete, Euclidean Distance Dendrogram" , cex=0.5,   horiz = FALSE)

```

```{r}

#pdf("h_clust_plot.pdf") 

dd <- dist(scale(wars_1500s), method = "euclidean")

hc <- hclust(dd, method = "complete")

dend <- hc %>% as.dendrogram() %>%
   set("branches_k_color", k=39) %>% 
  set("branches_lwd", 1)  %>%     
  set("labels_cex", c(.15,0.15)) 

# plot the dend in usual "base" plotting engine:
plot(dend,col="white", type="rectangle", frame.plot = FALSE, center	=FALSE, xlab="Belligerents", ylab="Distance (Euclidean)", main="Complete, Euclidean Distance Dendrogram" , cex=0.5,   horiz = FALSE)

```

```{r}

#pdf("h_clust_plot.pdf") 

dd <- dist(scale(wars_1600s), method = "euclidean")

hc <- hclust(dd, method = "complete")

dend <- hc %>% as.dendrogram() %>%
   set("branches_k_color", k=39) %>% 
  set("branches_lwd", 1)  %>%     
  set("labels_cex", c(.15,0.15)) 

# plot the dend in usual "base" plotting engine:
plot(dend,col="white", type="rectangle", frame.plot = FALSE, center	=FALSE, xlab="Belligerents", ylab="Distance (Euclidean)", main="Complete, Euclidean Distance Dendrogram" , cex=0.5,   horiz = FALSE)

```

```{r}

#pdf("h_clust_plot.pdf") 

dd <- dist(scale(wars_1700s), method = "euclidean")

hc <- hclust(dd, method = "complete")

dend <- hc %>% as.dendrogram() %>%
   set("branches_k_color", k=39) %>% 
  set("branches_lwd", 1)  %>%     
  set("labels_cex", c(.15,0.15)) 

# plot the dend in usual "base" plotting engine:
plot(dend,col="white", type="rectangle", frame.plot = FALSE, center	=FALSE, xlab="Belligerents", ylab="Distance (Euclidean)", main="Complete, Euclidean Distance Dendrogram" , cex=0.5,   horiz = FALSE)

```


```{r}

#pdf("h_clust_plot.pdf") 

set.seed(883)

dd <- dist(scale(wars_1800s), method = "euclidean")

hc <- hclust(dd, method = "complete")

dend <- hc %>% as.dendrogram() %>%
   set("branches_k_color", k=39) %>% 
  set("branches_lwd", 1)  %>%     
  set("labels_cex", c(.1,0.1)) 

# plot the dend in usual "base" plotting engine:
plot(dend,col="white", type="rectangle", frame.plot = FALSE, center	=FALSE, xlab="Belligerents", ylab="Distance (Euclidean)", main="Complete, Euclidean Distance Dendrogram" , cex=0.5,   horiz = FALSE)

```

### Edge Prediction Analysis

#### Edge Prediction 1000s

Edge prediction will initally be done with predict_edges(), a function in igraph with stron influence from the work of Aaron Clauset, including "Hierarchical structure and the prediction of missing links in networks." This method of edge prediction gives the probability of an edge being missing between two nodes using HRG which is sampled in MCMC or Markov chain Monte Carlo. The MCMC characterizes the distribution of the data by randomly sampling. The first aspect of MCMC, Monte-Carlos "is the practice of estimating the properties of a distribution by examining random samples from the distribution."" (van Ravenzwaaij 143) The Latter part, Markov chain suggests that random samples are generated sequentially. This allows us to focus on a posterior distribution rather than assuming one in our data-generating-process.

Clauset explains the process as taking an input graph initially and runs MCMC, each time sampling a dendrogram with corresponding log-likelihoods. This MCMC then outputs a list of possible missing connections in descending order of log-likelihood. 

Next I will look to see if the simulations replicate some of the structure and connections of the 1100s data set. The 1100s data set has more edges than the 1000s so we will simulate conflicts to see if the simulations will result in wars that are seen in the 1100s. I will simulate a number of wars equivalent to the difference between the two. The 1100s have 238 edges and the 1000s have 153 edges. I will add these 85 edges.

```{r}
(238-153) %>% kable()
```

This is documented in the R Documentation for predict_edges 

A simple introduction to Markov Chain Monte-Carlo sampling.

"Structural Inference of Hierarchies in Networks"

van Ravenzwaaij, Don et al. A simple introduction to Markov Chain Monte-Carlo sampling. Psychonomic bulletin & review vol. 25,1 (2018): 143-154. doi:10.3758/s13423-016-1015-8

First I will start with 7 new edges that the simulations give a log-likelihood over 700

```{r}

set.seed(1287)

prediction_for_1000s <- predict_edges(
  wars_in_1000s.ig,
  hrg = NULL,
  start = FALSE,
  num.samples = 100,
  num.bins = 25
)

#Here I look for predicted edge probability over 33% as a test
prediction_for_1000s$prob[c(1:7)]
# It turns out that 28 predicted edges are given a probability over 33%
prediction_for_1000s_best_7 <- (prediction_for_1000s$edges)[c(1:7),]
# I then create a new rename of my graph object
wars_in_1000s_for_adding_predicted_edges.ig <- wars_in_1000s.ig

wars_in_1000s_added_edges.ig <- add_edges(wars_in_1000s_for_adding_predicted_edges.ig, 
                                       prediction_for_1000s_best_7)


E(wars_in_1000s_added_edges.ig)[1:153]$color = "gray50"

E(wars_in_1000s_added_edges.ig)[154:160]$color = "red"

# wars_in_1000s_added_edges_network<- asNetwork(wars_in_1000s_added_edges.ig)

```

```{r}

set.seed(1287)

colrs <- c("red", "gray50")

set.seed(777)
plot(
  wars_in_1000s_added_edges.ig, 
     vertex.color = "gray78",
     vertex.size=4, 
     edge.arrow.width=0.5, 
     edge.arrow.size=.2, 
     edge.curved=.2, 
     vertex.label.cex=0.2,  
     vertex.label.color="black"
     )

legend(x=-1.5, y=-1.1, c("Simulated Edge Color", "Original Edge"), pch=21,

       col="#777777", pt.bg=colrs, pt.cex=2, cex=.8, bty="n", ncol=1)

```
```{r}
# Circle layout

set.seed(1287)

l <- layout_in_circle(wars_in_1000s_added_edges.ig)

plot(wars_in_1000s_added_edges.ig, layout=l,    vertex.color = "gray78",
     vertex.size=4, 
     edge.arrow.width=0.5, 
     edge.arrow.size=.2, 
     edge.curved=.2, 
     vertex.label.cex=0.2,  
     vertex.label.color="black"
     )
legend(x=-1.5, y=-1.1, c("Simulated Edge Color", "Original Edge"), pch=21,

       col="#777777", pt.bg=colrs, pt.cex=2, cex=.8, bty="n", ncol=1)

```

```{r}

set.seed(1287)

# dh fr lgl
layout_dh_1000s <- layout_with_dh(wars_in_1000s_added_edges.ig)

plot(wars_in_1000s_added_edges.ig, layout=layout_dh_1000s,    
     vertex.color = "gray78",
     vertex.size=4, 
     edge.arrow.width=0.5, 
     edge.arrow.size=.2, 
     edge.curved=.2, 
     vertex.label.cex=0.2,  
     vertex.label.color="black"
     )
legend(x=-1.5, y=-1.1, c("Simulated Edge Color", "Original Edge"), pch=21,

       col="#777777", pt.bg=colrs, pt.cex=2, cex=.8, bty="n", ncol=1)

```

#### Comparison with 85 new edges

```{r}

set.seed(90234)

prediction_for_1000s_85 <- predict_edges(
  wars_in_1000s.ig,
  hrg = NULL,
  start = FALSE,
  num.samples = 100,
  num.bins = 25
)

#Here I look for predicted edge probability over 33% as a test

# It turns out that 28 predicted edges are given a probability over 33%
prediction_for_1000s_best_85 <- (prediction_for_1000s$edges)[c(1:85),]
# I then create a new rename of my graph object
wars_in_1000s_for_adding_predicted_edges_85.ig <- wars_in_1000s.ig

wars_in_1000s_added_edges_85.ig <- add_edges(wars_in_1000s_for_adding_predicted_edges_85.ig, 
                                       prediction_for_1000s_best_85)

E(wars_in_1000s_added_edges_85.ig)[1:153]$color = "gray50"

E(wars_in_1000s_added_edges_85.ig)[154:238]$color = "red"

# wars_in_1000s_added_edges_network<- asNetwork(wars_in_1000s_added_edges.ig)

```

```{r}
# dh fr lgl

set.seed(1287)

layout_dh_1000s_85 <- layout_with_dh(wars_in_1000s_added_edges_85.ig)

plot(wars_in_1000s_added_edges_85.ig, layout=layout_dh_1000s_85,    
     vertex.color = "gray78",
     vertex.size=4, 
     edge.arrow.width=0.5, 
     edge.arrow.size=.2, 
     edge.curved=.2, 
     vertex.label.cex=0.2,  
     vertex.label.color="black"
     )
legend(x=-1.5, y=-1.1, c("Simulated Edge Color", "Original Edge"), pch=21,

       col="#777777", pt.bg=colrs, pt.cex=2, cex=.8, bty="n", ncol=1)


```

Package NetworkComparisonTest

"This permutation based hypothesis test, suited for gaussian and binary data, assesses the difference between two networks based on several invariance measures (network structure invariance, global strength invariance, edge invariance). Network structures are estimated with l1-regularized partial correlations (gaussian data) or with l1-regularized logistic regression (eLasso, binary data). Suited for comparison of independent and dependent samples. For dependent samples, only supported for data of one group which is measured twice."

http://eliassi.org/papers/soundarajan-sdm14.pdf

```{r}

set.seed(1287)

wars_in_1000s_added_edges_7_network <- average.path.length(wars_in_1000s_added_edges.ig,directed=T)

wars_in_1000s_added_edges_network <- as.network(as.matrix(wars_in_1000s_added_edges.ig))

wars_in_1000s_added_edges_85_network <- as.network(as.matrix(wars_in_1000s_added_edges_85.ig))

trans_for_added_edges_global <- c(
  
transitivity(wars_in_1000s.ig,  type="global"),

transitivity(wars_in_1100s.ig,  type="global"),

transitivity(wars_in_1000s_added_edges.ig,  type="global"),

transitivity(wars_in_1000s_added_edges_85.ig,  type="global")

)

trans_for_added_edges_averge <- c(
  
transitivity(wars_in_1000s.ig,  type="average"),

transitivity(wars_in_1100s.ig,  type="average"),

transitivity(wars_in_1000s_added_edges.ig,  type="average"),

transitivity(wars_in_1000s_added_edges_85.ig,  type="average")

)

apl_for_added_edges <- c(

average.path.length(wars_in_1000s.ig,directed=T),

average.path.length(wars_in_1100s.ig,directed=T),

average.path.length(wars_in_1000s_added_edges.ig,directed=T),

average.path.length(wars_in_1000s_added_edges_85.ig,directed=T)

)

network_density_for_added_edges <- c(

network.density(wars_in_1000s_network),

network.density(wars_in_1100s_network),

network.density(wars_in_1000s_added_edges_network),

network.density(wars_in_1000s_added_edges_85_network)

)

degree_centralization_for_added_edges <- c(
centr_degree(wars_in_1000s.ig, loops = FALSE, mode="in")$centralization,

centr_degree(wars_in_1100s.ig, loops = FALSE, mode="in")$centralization,

centr_degree(wars_in_1000s_added_edges.ig, loops = FALSE, mode="in")$centralization,

centr_degree(wars_in_1000s_added_edges_85.ig, loops = FALSE, mode="in")$centralization

)
data_frame(
  c(trans_for_added_edges_global, 
  trans_for_added_edges_averge, 
  apl_for_added_edges, 
  network_density_for_added_edges, 
  degree_centralization_for_added_edges
  )
)

groups_for_added_edges <- c("Wars in 1000s", "Wars in 1100s", "Wars in 1000s with 7 Edges", "Wars in 1000s with 85 Edges")

df_for_edge_add_comp <- data.frame(trans_for_added_edges_global, 
  trans_for_added_edges_averge, 
  apl_for_added_edges, 
  network_density_for_added_edges, 
  degree_centralization_for_added_edges,groups_for_added_edges, row.names=c("Wars in 1000s", "Wars in 1100s", "Wars in 1000s with 7 Edges", "Wars in 1000s with 85 Edges"))

names(df_for_edge_add_comp)[1:5] <- c("Global Transitivity", "Average Transitivity", "Average Path Length", "Network Density", "Degree Centralization")
df_for_edge_add_comp

```


#### Edge Prediction1100s

```{r}

set.seed(1287)

ggplot(df_for_edge_add_comp, aes(x=`Global Transitivity`, fill=groups_for_added_edges)) + 
  geom_histogram() +
  scale_fill_manual(values = c("Wars in 1000s" = "#CD7F32",
                               "Wars in 1100s" = "#C0C0C0",
                               "Wars in 1000s with 7 Edges" = "gold",
                               "Wars in 1000s with 85 Edges" = "red"))

ggplot(df_for_edge_add_comp, aes(x=`Average Transitivity`, fill=groups_for_added_edges)) + 
  geom_histogram() +
  scale_fill_manual(values = c("Wars in 1000s" = "#CD7F32",
                               "Wars in 1100s" = "#C0C0C0",
                               "Wars in 1000s with 7 Edges" = "gold",
                               "Wars in 1000s with 85 Edges" = "red"))

ggplot(df_for_edge_add_comp, aes(x=`Average Path Length`, fill=groups_for_added_edges)) + 
  geom_histogram() +
  scale_fill_manual(values = c("Wars in 1000s" = "#CD7F32",
                               "Wars in 1100s" = "#C0C0C0",
                               "Wars in 1000s with 7 Edges" = "gold",
                               "Wars in 1000s with 85 Edges" = "red"))
ggplot(df_for_edge_add_comp, aes(x=`Network Density`, fill=groups_for_added_edges)) + 
  geom_histogram() +
  scale_fill_manual(values = c("Wars in 1000s" = "#CD7F32",
                               "Wars in 1100s" = "#C0C0C0",
                               "Wars in 1000s with 7 Edges" = "gold",
                               "Wars in 1000s with 85 Edges" = "red"))

ggplot(df_for_edge_add_comp, aes(x=`Degree Centralization`, fill=groups_for_added_edges)) + 
  geom_histogram() +
  scale_fill_manual(values = c("Wars in 1000s" = "#CD7F32",
                               "Wars in 1100s" = "#C0C0C0",
                               "Wars in 1000s with 7 Edges" = "gold",
                               "Wars in 1000s with 85 Edges" = "red"))


```

```{r}

set.seed(1287)

u1 <- as.undirected(wars_in_1000s.ig)
u2 <- as.undirected(wars_in_1100s.ig)
u3 <- as.undirected(wars_in_1000s_added_edges.ig)
u4 <- as.undirected(wars_in_1000s_added_edges_85.ig)


same_1=length(E(intersection(wars_in_1000s.ig, wars_in_1100s.ig)))

difference_1 <- length(E(difference(wars_in_1000s.ig, wars_in_1100s.ig)))

reverse_1=length(E(difference(wars_in_1000s.ig,wars_in_1100s.ig )))-length(E(difference(u2,u1)))

c(same_1, difference_1, reverse_1)

same_2=length(E(intersection(wars_in_1100s.ig, wars_in_1000s_added_edges.ig)))

difference_2 <- length(E(difference(wars_in_1100s.ig, wars_in_1000s_added_edges.ig)))

c(same_2, difference_2)

same_3=length(E(intersection(wars_in_1100s.ig, wars_in_1000s_added_edges_85.ig)))

difference_3 <- length(E(difference(wars_in_1100s.ig, wars_in_1000s_added_edges_85.ig)))

c(same_3, difference_3)

reverse_1=length(E(difference(wars_in_1000s.ig,wars_in_1100s.ig)))-length(E(difference(u2,u1)))

addition_1=length(E(difference(wars_in_1000s.ig,wars_in_1100s.ig)))-reverse_1

same_2 = length(E(intersection(wars_in_1100s.ig, wars_in_1000s_added_edges.ig)))

reverse_2=length(E(difference(wars_in_1100s.ig, wars_in_1000s_added_edges.ig)))-length(E(difference(u2,u3)))

addition_2=length(E(difference(wars_in_1100s.ig, wars_in_1000s_added_edges.ig)))-reverse_2

same_3 = length(E(intersection(wars_in_1100s.ig, wars_in_1000s_added_edges_85.ig)))


difference_3 = length(E(difference(wars_in_1100s.ig, wars_in_1000s_added_edges_85.ig)))
       
reverse_3 = length(E(difference(wars_in_1100s.ig, wars_in_1000s_added_edges_85.ig)))-length(E(difference(u2,u4)))

addition_3 = length(E(difference(wars_in_1100s.ig, wars_in_1000s_added_edges_85.ig)))-reverse_2

```

#### Edge Prediction 1100s

```{r}

set.seed(1287)

prediction_for_1100s <- predict_edges(
  wars_in_1100s.ig,
  hrg = NULL,
  start = FALSE,
  num.samples = 100,
  num.bins = 25
)

#Here I look for predicted edge probability over 33% as a test
prediction_for_1100s$prob[c(1:7)]
# It turns out that 29 predicted edges are given a probability over 33%
prediction_for_1100s_best_7 <- (prediction_for_1100s$edges)[c(1:7),]
# I then create a new rename of my graph object
wars_in_1100s_for_adding_predicted_edges.ig <- wars_in_1100s.ig

wars_in_1100s_added_edges.ig <- add_edges(wars_in_1100s_for_adding_predicted_edges.ig, 
                                       prediction_for_1100s_best_7)


E(wars_in_1100s_added_edges.ig)[1:153]$color = "gray50"

E(wars_in_1100s_added_edges.ig)[154:160]$color = "red"

# wars_in_1000s_added_edges_network<- asNetwork(wars_in_1000s_added_edges.ig)

```

```{r}

set.seed(90234)

prediction_for_1100s_75 <- predict_edges(
  wars_in_1100s.ig,
  hrg = NULL,
  start = FALSE,
  num.samples = 100,
  num.bins = 25
)

#Here I look for predicted edge probability over 33% as a test

# It turns out that 28 predicted edges are given a probability over 33%
prediction_for_1100s_best_75 <- (prediction_for_1100s$edges)[c(1:75),]
# I then create a new rename of my graph object
wars_in_1100s_for_adding_predicted_edges_75.ig <- wars_in_1100s.ig

wars_in_1100s_added_edges_75.ig <- add_edges(wars_in_1100s_for_adding_predicted_edges_75.ig, 
                                       prediction_for_1100s_best_75)

E(wars_in_1100s_added_edges_75.ig)[1:238]$color = "gray50"

E(wars_in_1100s_added_edges_75.ig)[239:313]$color = "red"

# wars_in_1000s_added_edges_network<- asNetwork(wars_in_1000s_added_edges.ig)

```

```{r}

set.seed(1287)

colrs <- c("red", "gray50")

set.seed(777)
plot(
  wars_in_1100s_added_edges_75.ig, 
     vertex.color = "gray78",
     vertex.size=4, 
     edge.arrow.width=0.5, 
     edge.arrow.size=.2, 
     edge.curved=.2, 
     vertex.label.cex=0.2,  
     vertex.label.color="black"
     )

legend(x=-1.5, y=-1.1, c("Simulated Edge Color", "Original Edge"), pch=21,

       col="#777777", pt.bg=colrs, pt.cex=2, cex=.8, bty="n", ncol=1)

```

```{r}

set.seed(1287)

undirected_1100s <- as.undirected(wars_in_1100s.ig)

undirected_1200s <- as.undirected(wars_in_1200s.ig)

undirected_1100s_added_edges <- as.undirected(wars_in_1100s_added_edges_75.ig)

intersection_1100s_1200s = length(E(intersection(undirected_1100s, undirected_1200s)))

difference_1100s_1200s <- length(E(difference(undirected_1100s, undirected_1200s)))

c(intersection_1100s_1200s, difference_1100s_1200s)

intersection_1200s_1100s_added_edges = length(E(intersection(undirected_1200s, undirected_1100s_added_edges)))

difference_1200s_1100s_added_edges <- length(E(difference(undirected_1200s, undirected_1100s_added_edges)))

c(intersection_1200s_1100s_added_edges, difference_1200s_1100s_added_edges)

```

#### Edge Prediction 1200s

This section is somewhat peculiar so it will be skipped, there was relatively less conflict going on, which one could likely attribute to the black plague. In fact the number of notable conflicts between my wars_in_1200s.ig and wars_in_1300s.ig decreases from 313 to 241. As a result it would be redundant to predict edges when not predicting the proliferation of conflict, but its decline due to a disease that crippled a center of worldwide conflict at the time.

#### Edge Prediction 1300s to 1400s

Fortunately in the 1300s we have much more room to predict for the 1400s as the period that followed the plague saw an escalation in conflict simply due to increasing population and nations simply having the ability to go to war once again.

##### predict_edges for 1300s to 1400s

```{r}

set.seed(1287)

prediction_for_1300s <- predict_edges(
  wars_in_1300s.ig,
  hrg = NULL,
  start = FALSE,
  num.samples = 1000,
  num.bins = 25
)

#Here I look at the 7 predicted edges with the highest probabilities
prediction_for_1300s$prob[c(1:7)]
# The difference in number of edges between the 1300s and 1400s is 85 so 
# I will predict 85 edges
prediction_for_1300s_best_85 <- (prediction_for_1300s$edges)[c(1:85),]
# I then create a new rename of my graph object
wars_in_1300s_for_adding_predicted_edges.ig <- wars_in_1300s.ig

wars_in_1300s_added_edges.ig <- add_edges(wars_in_1300s_for_adding_predicted_edges.ig, 
                                       prediction_for_1300s_best_85)


E(wars_in_1300s_added_edges.ig)[1:241]$color = "gray50"

E(wars_in_1300s_added_edges.ig)[242:326]$color = "red"

# wars_in_1000s_added_edges_network<- asNetwork(wars_in_1000s_added_edges.ig)

```

```{r}

set.seed(1287)

undirected_1300s <- as.undirected(wars_in_1300s.ig)

undirected_1400s <- as.undirected(wars_in_1400s.ig)

undirected_1300s_added_edges <- as.undirected(wars_in_1300s_added_edges.ig)

intersection_1300s_1400s = length(E(intersection(undirected_1300s, undirected_1400s)))

difference_1300s_1400s <- length(E(difference(undirected_1300s, undirected_1400s)))

c(intersection_1300s_1400s, difference_1300s_1400s)

intersection_1400s_1300s_added_edges = length(E(intersection(undirected_1400s, undirected_1300s_added_edges)))

difference_1400s_1300s_added_edges <- length(E(difference(undirected_1400s, undirected_1300s_added_edges)))

c(intersection_1400s_1300s_added_edges, difference_1400s_1300s_added_edges)

```

#### Edge Prediction 1400s to 1500s

In the prior section we were correctly able to predict no more than the typical number of times. However, I will add a subsequent section that **removes** directness before edge calculation. This is because I noticed that adding predicted edges to the directed networks resulted in them being removed when making it undirected, thus I will make the network undirected first before analysis. I admit this likely loses some significant predictive implications. However, we will simply being seeing **if** a conflict occurs rather than predicting the conflct __and__ who wins.

##### predict_edges for 1400s to 1500s

```{r}

set.seed(1287)

prediction_for_1400s <- predict_edges(
  wars_in_1400s.ig,
  hrg = NULL,
  start = FALSE,
  num.samples = 1000,
  num.bins = 25
)

#Here I look at the 7 predicted edges with the highest probabilities
prediction_for_1400s$prob[c(1:7)]
# The difference in number of edges between the 1300s and 1400s is 85 so 
# I will predict 85 edges
prediction_for_1400s_best_419 <- (prediction_for_1400s$edges)[c(1:419),]
# I then create a new rename of my graph object
wars_in_1400s_for_adding_predicted_edges.ig <- wars_in_1400s.ig

wars_in_1400s_added_edges.ig <- add_edges(wars_in_1400s_for_adding_predicted_edges.ig, 
                                       prediction_for_1400s_best_419)


E(wars_in_1400s_added_edges.ig)[1:326]$color = "gray50"

E(wars_in_1400s_added_edges.ig)[327:745]$color = "red"

# wars_in_1000s_added_edges_network<- asNetwork(wars_in_1000s_added_edges.ig)


```

```{r}

set.seed(1287)

undirected_1400s <- as.undirected(wars_in_1400s.ig)

undirected_1500s <- as.undirected(wars_in_1500s.ig)

undirected_1400s_added_edges <- as.undirected(wars_in_1400s_added_edges.ig)

intersection_1400s_1500s = length(E(intersection(undirected_1400s, undirected_1500s)))

difference_1400s_1500s <- length(E(difference(undirected_1400s, undirected_1500s)))

c(intersection_1400s_1500s, difference_1400s_1500s)

intersection_1500s_1400s_added_edges = length(E(intersection(undirected_1500s, undirected_1400s_added_edges)))

difference_1500s_1400s_added_edges <- length(E(difference(undirected_1500s, undirected_1400s_added_edges)))

c(intersection_1500s_1400s_added_edges, difference_1500s_1400s_added_edges)

```

### Edge Prediction in an Undirected Network.

#### predict_edges for 1400s to 1500s Undirected

```{r}

set.seed(1287)

# undirected_1400s <- as.undirected(wars_in_1400s.ig)

# 301 edges

# undirected_1500s <- as.undirected(wars_in_1500s.ig)

# 603 edges

# 603-301 = 302

prediction_for_1400s_undirected <- predict_edges(
  undirected_1400s,
  hrg = NULL,
  start = FALSE,
  num.samples = 1000,
  num.bins = 100
)

#Here I look at the 7 predicted edges with the highest probabilities
#prediction_for_1400s_undirected$prob[c(1:7)]

# The difference in number of edges between the 1400s and 1500s is 302 so 
# I will predict 302 edges in the undirected network
prediction_for_1400s_undirected_best_302 <- (prediction_for_1400s_undirected$edges)[c(1:302),]
# I then create a new rename of my graph object
wars_in_1400s_for_adding_edges_undirected.ig <- undirected_1400s

wars_in_1400s_added_edges_undirected.ig <- add_edges(wars_in_1400s_for_adding_edges_undirected.ig, 
                                       prediction_for_1400s_undirected_best_302)


E(wars_in_1400s_added_edges_undirected.ig)[1:301]$color = "gray50"

E(wars_in_1400s_added_edges_undirected.ig)[302:603]$color = "red"

# wars_in_1000s_added_edges_network<- asNetwork(wars_in_1000s_added_edges.ig)


```

```{r}

set.seed(1287)

# undirected_1400s <- as.undirected(wars_in_1400s.ig)

# undirected_1500s <- as.undirected(wars_in_1500s.ig)

#wars_in_1400s_added_edges_undirected.ig

intersection_1400s_1500s = length(E(intersection(undirected_1400s, undirected_1500s)))

difference_1400s_1500s <- length(E(difference(undirected_1400s, undirected_1500s)))

c(intersection_1400s_1500s, difference_1400s_1500s)

intersection_1500s_1400s_added_edges = length(E(intersection(undirected_1500s, wars_in_1400s_added_edges_undirected.ig)))

difference_1500s_1400s_added_edges <- length(E(difference(undirected_1500s, wars_in_1400s_added_edges_undirected.ig)))

c(intersection_1500s_1400s_added_edges, difference_1500s_1400s_added_edges)

```

#### predict_edges for 1600s to 1700s Undirected

```{r}

set.seed(1287)

undirected_1600s <- as.undirected(wars_in_1600s.ig)

# undirected_1500s

# 598 edges

undirected_1700s <- as.undirected(wars_in_1700s.ig)

# 1027 edges

# 1027-598 = 302

prediction_for_1600s_undirected <- predict_edges(
  undirected_1600s,
  hrg = NULL,
  start = FALSE,
  num.samples = 1000,
  num.bins = 100
)

#Here I look at the 7 predicted edges with the highest probabilities
#prediction_for_1400s_undirected$prob[c(1:7)]

# The difference in number of edges between the 1400s and 1500s is 429 so 
# I will predict 302 edges in the undirected network

prediction_for_1600s_undirected_best_429 <- (prediction_for_1600s_undirected$edges)[c(1:429),]
# I then create a new rename of my graph object

wars_in_1600s_for_adding_edges_undirected.ig <- undirected_1600s

wars_in_1600s_added_edges_undirected.ig <- add_edges(wars_in_1600s_for_adding_edges_undirected.ig, 
                                       prediction_for_1600s_undirected_best_429)


E(wars_in_1600s_added_edges_undirected.ig)[1:429]$color = "gray50"

E(wars_in_1600s_added_edges_undirected.ig)[430:1027]$color = "red"

# wars_in_1000s_added_edges_network<- asNetwork(wars_in_1000s_added_edges.ig)


```

```{r}

set.seed(1287)

# undirected_1600s <- as.undirected(wars_in_1400s.ig)

# undirected_1700s <- as.undirected(wars_in_1500s.ig)

#wars_in_1600s_added_edges_undirected.ig

intersection_1600s_1700s = length(E(intersection(undirected_1600s, undirected_1700s)))

difference_1600s_1700s <- length(E(difference(undirected_1600s, undirected_1700s)))

c(intersection_1600s_1700s, difference_1600s_1700s)

intersection_1700s_1600s_added_edges = length(E(intersection(undirected_1700s, wars_in_1600s_added_edges_undirected.ig)))

difference_1700s_1600s_added_edges <- length(E(difference(undirected_1700s, wars_in_1600s_added_edges_undirected.ig)))

c(intersection_1700s_1600s_added_edges, difference_1700s_1600s_added_edges)

```

##### Analysis of intersection for 1600s Simulated Nodes on 1700s

As can be seen above the this method only correctly predicts 2 additional edges, those being conflicts between Denmark-Norway and the Dutch Republic and Denmark-Norway and the Holy Roman Empire.

```{r}

difference(
  intersection(undirected_1700s, wars_in_1600s_added_edges_undirected.ig), intersection(undirected_1600s, undirected_1700s)
  )

```

to see what the probabilities of these were we must look into our original, undirected igraph data.

### Mapped Data 1800s

```{r}
library('maps')
library('geosphere')
library(diagram)
library(plotrix)
```

```{r, warning=FALSE}

pdf("1800_map.pdf") 
set.seed(15887) 
aspects_of_1800s_states_for_map <- read_excel("~/Desktop/Spring 2022/Networks/aspects_of_1800s_states_for_map.xlsx")

aspects_of_1800s_states_for_map <- aspects_of_1800s_states_for_map[,c(1:3)]

cl <- colors(distinct = TRUE)

# to set random generator seed

mycols2 <- sample(cl, 502)
mycols3 <- sample(cl, 145)
mycols2<-c(mycols2, mycols3)
palette(mycols2)

map('world', fill = TRUE, col = "white", mar = rep(0, 4), resolution = 10 , myborder = 0.001) #bg="skyblue1"

nodes <- transform(aspects_of_1800s_states_for_map, countries = as.factor(aspects_of_1800s_states_for_map$name))

with(nodes, points(long, Lat, col=countries, pch=19, cex=0.25)) 

edges <- subset(data.frame(from = wars_in_1800s$In, to = wars_in_1800s$Out, stringsAsFactors = F), from != to)

edges <- merge(merge(edges, nodes[, c("name", "long", "Lat")], by.x = "from", by.y = "name"), nodes[, c("name", "long", "Lat")], by.x = "to", by.y = "name")

edges$col <- as.integer(nodes$countries[match(edges$from, nodes$name)])

edges$long.x<-as.integer(edges$long.x)

edges$Lat.x<-as.integer(edges$Lat.x)

edges$long.y<-as.integer(edges$long.y)

edges$Lat.y<-as.integer(edges$Lat.y)

apply(edges[,-(1:2)], 1, function(x)
curvedarrow(to=x[3:4], from=x[1:2], lcol=x[5], 
            curve=.1, arr.pos = 1, lwd=.125, arr.lwd=0.01, 
                          arr.width=0.01, arr.length=0.1))

```


### Turning to Different Predictive Methods

https://igraph.org/r/doc/

### Back to Nodes

From here I will continue my analysis analysis on the nodes in my network, rather than continuing edge prediction.

```{r}

nodes_1000s$proportion_of_wins <- (
nodes_1000s$outdegree / nodes_1000s$degree
)
  
nodes_1100s$proportion_of_wins <- (
nodes_1100s$outdegree / nodes_1100s$degree
)

nodes_1200s$proportion_of_wins <- (
nodes_1200s$outdegree / nodes_1200s$degree
)

nodes_1300s$proportion_of_wins <- (
nodes_1300s$outdegree / nodes_1300s$degree
)

nodes_1400s$proportion_of_wins <- (
nodes_1400s$outdegree / nodes_1400s$degree
)

nodes_1500s$proportion_of_wins <- (
nodes_1500s$outdegree / nodes_1500s$degree
)

nodes_1600s$proportion_of_wins <- (
nodes_1600s$outdegree / nodes_1600s$degree
)

nodes_1700s$proportion_of_wins <- (
nodes_1700s$outdegree / nodes_1700s$degree
)

nodes_1800s$proportion_of_wins <- (
nodes_1800s$outdegree / nodes_1800s$degree
)

nodes_list <- list(nodes_1000s, nodes_1100s, nodes_1200s, nodes_1300s, 
                  nodes_1400s, nodes_1500s, nodes_1600s, nodes_1700s, 
                  nodes_1800s)

function_for_appending_nodes_with_prop_of_wins  <- function(x) { 
      x %>%       
          mutate(proportion_of_wins <- outdegree / degree)
}

function_for_appending_nodes_with_prop_of_losses  <- function(x) { 
      x %>%       
          mutate(proportion_of_losses <- indegree / degree)
}

nodes_1000s <- function_for_appending_nodes_with_prop_of_wins(nodes_1000s)

nodes_1100s <- function_for_appending_nodes_with_prop_of_wins(nodes_1100s)

nodes_1200s <- function_for_appending_nodes_with_prop_of_wins(nodes_1200s)

nodes_1300s <- function_for_appending_nodes_with_prop_of_wins(nodes_1300s)

nodes_1400s <- function_for_appending_nodes_with_prop_of_wins(nodes_1400s)

nodes_1500s <- function_for_appending_nodes_with_prop_of_wins(nodes_1500s)

nodes_1600s <- function_for_appending_nodes_with_prop_of_wins(nodes_1600s)

nodes_1700s <- function_for_appending_nodes_with_prop_of_wins(nodes_1700s)

nodes_1800s <- function_for_appending_nodes_with_prop_of_wins(nodes_1800s)

```

```{r}

nodes_1000s %>% select(outdegree, name) %>%
ggplot(aes(x=reorder(name, -outdegree), y=outdegree))+ 
  geom_bar(position = 'dodge', stat= 'identity', aes(fill=name), show.legend = FALSE) + 
  #geom_text(aes(label = name), size = 1, vjust = -0.5) +
   labs(title = "Yearly Expenses by Category",
        y = "Total Expense (USD)",
        x = "Primary Expense Category") + 
  theme_minimal(base_size=1,  base_family="Georgia")

```



