---
title: "Networks Blog Post 8"
description: |
  This blog post covers material from week 9 of Social Networks.
author:
  - name: Noah Milstein
    url: {}
date: 2022-04-05
output:
  distill::distill_article:
    self_contained: false
---

```{r, echo=FALSE}
library(readxl)
library(igraph)
library(statnet)
library(network)
library(tidyverse)
library(ggnetwork)
library(GGally)
library(ggplot2)
library(sna)
library(intergraph)
library(ggbiplot)
library(knitr)
library(GGally)
library(aRtsy)
```

# Loading the Networks


```{r setup, include=FALSE}

wars_in_1000s <- read_excel("~/Desktop/Spring 2022/Networks/wars_in_1000s.xlsx")

wars_in_1000s$"Out-list" <- sapply(wars_in_1000s$"Out-list",
                                    function(x) { gsub("[\r\n]", "", x) })
wars_in_1000s$"In-list"<- sapply(wars_in_1000s$"In-list",
                                    function(x) { gsub("[\r\n]", "", x) })

wars_in_1100s <- read_excel("~/Desktop/Spring 2022/Networks/wars_in_1100s.xlsx")

wars_in_1100s$"Out" <- sapply(wars_in_1100s$"Out",
                                    function(x) { gsub("[\r\n]", "", x) })
wars_in_1100s$"In"<- sapply(wars_in_1100s$"In",
                                    function(x) { gsub("[\r\n]", "", x) })

wars_in_1200s <- read_excel("~/Desktop/Spring 2022/Networks/wars_in_1200s.xlsx")


wars_in_1200s$"Out" <- sapply(wars_in_1200s$"Out",
                                    function(x) { gsub("[\r\n]", "", x) })
wars_in_1200s$"In"<- sapply(wars_in_1200s$"In",  
                                    function(x) { gsub("[\r\n]", "", x) })
                            
```

## Putting Network into Necessary Formats

```{r, echo=FALSE}

wars_in_1000s_edgelist <- as.matrix(wars_in_1000s)

wars_in_1000s_edgelist_network_edgelist <- graph.edgelist(wars_in_1000s_edgelist, directed=TRUE)

wars_in_1000s.ig<-graph_from_data_frame(wars_in_1000s)

wars_in_1000s_network <- asNetwork(wars_in_1000s.ig)

wars_in_1100s_edgelist <- as.matrix(wars_in_1100s)

wars_in_1100s_edgelist_network_edgelist <- graph.edgelist(wars_in_1100s_edgelist, directed=TRUE)

wars_in_1100s.ig<-graph_from_data_frame(wars_in_1100s)

wars_in_1100s_network <- asNetwork(wars_in_1100s.ig)

wars_in_1200s_edgelist <- as.matrix(wars_in_1200s)

wars_in_1200s_edgelist_network_edgelist <- graph.edgelist(wars_in_1200s_edgelist, directed=TRUE)

wars_in_1200s.ig<-graph_from_data_frame(wars_in_1200s)

wars_in_1200s_network <- asNetwork(wars_in_1200s.ig)

```

## Adding Attributes

```{r,echo=FALSE}

wars_in_1000s.nodes.stat<-data.frame(name=wars_in_1000s_network%v%"vertex.names",
    totdegree=sna::degree(wars_in_1000s_network),
    indegree=sna::degree(wars_in_1000s_network, cmode="indegree"),
    outdegree=sna::degree(wars_in_1000s_network, cmode="outdegree"))

wars_in_1000s.nodes.stat$eigen<-evcent(wars_in_1000s_network)

wars_in_1000s.nodes.stat$close<-sna::closeness(wars_in_1000s_network, cmode="suminvdir")

wars_1000s<-as.matrix(as_adjacency_matrix(wars_in_1000s.ig))

#square the adjacency matrix

wars_1000s_sq<-t(wars_1000s) %*% wars_1000s

#Calculate the proportion of reflected centrality.

wars_in_1000s.nodes.stat$rc<-diag(wars_1000s_sq)/rowSums(wars_1000s_sq)

#replace missing values with 0

wars_in_1000s.nodes.stat$rc<-ifelse(is.nan(wars_in_1000s.nodes.stat$rc),0,wars_in_1000s.nodes.stat$rc)

#Calculate received eigenvalue centrality
wars_in_1000s.nodes.stat$eigen.rc<-wars_in_1000s.nodes.stat$eigen*wars_in_1000s.nodes.stat$rc

#Calculate the proportion of derived centrality.
wars_in_1000s.nodes.stat$dc<-1-diag(wars_1000s_sq)/rowSums(wars_1000s_sq)
#replace missing values with 0
wars_in_1000s.nodes.stat$dc<-ifelse(is.nan(wars_in_1000s.nodes.stat$dc),1,wars_in_1000s.nodes.stat$dc)
#Calculate received eigenvalue centrality
wars_in_1000s.nodes.stat$eigen.dc<-wars_in_1000s.nodes.stat$eigen*wars_in_1000s.nodes.stat$dc

```

#### Brokerage scores in the 1000s

```{r,echo=FALSE}

temp<-data.frame(brokerage(wars_in_1000s_network, cl = wars_in_1000s.nodes.stat$totdegree)$z.nli)

wars_in_1000s.nodes.stat_2<-wars_in_1000s.nodes.stat %>%
  mutate(broker.tot = temp$t,
         broker.coord = temp$w_I,
         broker.itin = temp$w_O,
         broker.rep = temp$b_IO,
         broker.gate = temp$b_OI,
         broker.lia = temp$b_O)

```


```{r, echo=FALSE}

wars_in_1100s.nodes.stat<-data.frame(name=wars_in_1100s_network%v%"vertex.names",
    totdegree=sna::degree(wars_in_1100s_network),
    indegree=sna::degree(wars_in_1100s_network, cmode="indegree"),
    outdegree=sna::degree(wars_in_1100s_network, cmode="outdegree"))

wars_in_1100s.nodes.stat$eigen<-evcent(wars_in_1100s_network)

wars_1100s<-as.matrix(as_adjacency_matrix(wars_in_1100s.ig))

#square the adjacency matrix

wars_1100s_sq<-t(wars_1100s) %*% wars_1100s

#Calculate the proportion of reflected centrality.

wars_in_1100s.nodes.stat$rc<-diag(wars_1100s_sq)/rowSums(wars_1100s_sq)

#replace missing values with 0

wars_in_1100s.nodes.stat$rc<-ifelse(is.nan(wars_in_1100s.nodes.stat$rc),0,wars_in_1100s.nodes.stat$rc)

#Calculate received eigenvalue centrality
wars_in_1100s.nodes.stat$eigen.rc<-wars_in_1100s.nodes.stat$eigen*wars_in_1100s.nodes.stat$rc

#Calculate the proportion of derived centrality.
wars_in_1100s.nodes.stat$dc<-1-diag(wars_1100s_sq)/rowSums(wars_1100s_sq)
#replace missing values with 0
wars_in_1100s.nodes.stat$dc<-ifelse(is.nan(wars_in_1100s.nodes.stat$dc),1,wars_in_1100s.nodes.stat$dc)
#Calculate received eigenvalue centrality
wars_in_1100s.nodes.stat$eigen.dc<-wars_in_1100s.nodes.stat$eigen*wars_in_1100s.nodes.stat$dc

```

#### Brokerage scores in the 1100s

```{r, echo=FALSE}

temp<-data.frame(brokerage(wars_in_1100s_network, cl = wars_in_1100s.nodes.stat$totdegree)$z.nli)

wars_in_1100s.nodes.stat_2<-wars_in_1100s.nodes.stat %>%
  mutate(broker.tot = temp$t,
         broker.coord = temp$w_I,
         broker.itin = temp$w_O,
         broker.rep = temp$b_IO,
         broker.gate = temp$b_OI,
         broker.lia = temp$b_O)

```

#### Brokerage scores in the 1200s

```{r, echo=FALSE}

wars_in_1200s.nodes.stat<-data.frame(name=wars_in_1200s_network%v%"vertex.names",
    totdegree=sna::degree(wars_in_1200s_network),
    indegree=sna::degree(wars_in_1200s_network, cmode="indegree"),
    outdegree=sna::degree(wars_in_1200s_network, cmode="outdegree"))

wars_in_1200s.nodes.stat$eigen<-evcent(wars_in_1200s_network)

```

```{r, echo=FALSE}

wars_1200s<-as.matrix(as_adjacency_matrix(wars_in_1200s.ig))

#square the adjacency matrix

wars_1200s_sq<-t(wars_1200s) %*% wars_1200s

#Calculate the proportion of reflected centrality.

wars_in_1200s.nodes.stat$rc<-diag(wars_1200s_sq)/rowSums(wars_1200s_sq)

#replace missing values with 0

wars_in_1200s.nodes.stat$rc<-ifelse(is.nan(wars_in_1200s.nodes.stat$rc),0,wars_in_1200s.nodes.stat$rc)

#Calculate received eigenvalue centrality
wars_in_1200s.nodes.stat$eigen.rc<-wars_in_1200s.nodes.stat$eigen*wars_in_1200s.nodes.stat$rc

#Calculate the proportion of derived centrality.
wars_in_1200s.nodes.stat$dc<-1-diag(wars_1200s_sq)/rowSums(wars_1200s_sq)
#replace missing values with 0
wars_in_1200s.nodes.stat$dc<-ifelse(is.nan(wars_in_1200s.nodes.stat$dc),1,wars_in_1200s.nodes.stat$dc)
#Calculate received eigenvalue centrality
wars_in_1200s.nodes.stat$eigen.dc<-wars_in_1200s.nodes.stat$eigen*wars_in_1200s.nodes.stat$dc

```

```{r, echo=FALSE}

temp<-data.frame(brokerage(wars_in_1200s_network, cl = wars_in_1200s.nodes.stat$totdegree)$z.nli)

wars_in_1200s.nodes.stat_2<-wars_in_1200s.nodes.stat %>%
  mutate(broker.tot = temp$t,
         broker.coord = temp$w_I,
         broker.itin = temp$w_O,
         broker.rep = temp$b_IO,
         broker.gate = temp$b_OI,
         broker.lia = temp$b_O)

```

# Inital Graphical Representation

### 1000s GGplot

```{r, echo=FALSE, warning=FALSE}

set.seed(777)
ggplot(wars_in_1000s_network, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "grey25", alpha = 0.25, arrow = arrow(length = unit(2, "pt"), type = "closed"))+
  geom_nodes(color = "blue", alpha = 0.3, size = 2) +
  geom_nodetext(aes( label = vertex.names ), size=1)+
    ggtitle("Networks of Wars in the 1000s") +
  theme_blank()

```

### 1100s GGplot

```{r, echo=FALSE, warning=FALSE}

set.seed(777)
ggplot(wars_in_1100s_network, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "grey25", alpha = 0.25, arrow = arrow(length = unit(2, "pt"), type = "closed"))+
  geom_nodes(color = "peru", alpha = 0.3, size = 2) +
  geom_nodetext(aes( label = vertex.names ), size=1)+
    ggtitle("Networks of Wars in the 1100s") +
  theme_blank()

```

### 1200s GGplot

```{r, echo=FALSE, warning=FALSE}
set.seed(777)
ggplot(wars_in_1200s_network, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "grey25", alpha = 0.25, arrow = arrow(length = unit(2, "pt"), type = "closed"))+
  geom_nodes(color = "lightpink", alpha = 0.3, size = 2) +
  geom_nodetext(aes( label = vertex.names ), size=1)+
    ggtitle("Networks of Wars in the 1200s") +
  theme_blank()
```

```{r}

wars_in_1000s_edgelist <- as.matrix(wars_in_1000s)

wars_in_1000s_edgelist_network_edgelist <- graph.edgelist(wars_in_1000s_edgelist, directed=TRUE)

wars_in_1000s.ig<-graph_from_data_frame(wars_in_1000s)

wars_in_1000s_network <- asNetwork(wars_in_1000s.ig)

```

```{r}


aspects_of_1000s_states <- read_excel("~/Desktop/Spring 2022/Networks/aspects_of_1000s_states.xlsx")

total_1000s <- merge(aspects_of_1000s_states, wars_in_1000s.nodes.stat_2, by="name")

```

```{r}

aspects_of_1100s_states <- read_excel("~/Desktop/Spring 2022/Networks/aspects_of_1100s_states.xlsx")

total_1100s <- merge(aspects_of_1100s_states, wars_in_1100s.nodes.stat_2, by="name")

```

```{r}

aspects_of_1200s_states <- read_excel("~/Desktop/Spring 2022/Networks/aspects_of_1200s_states.xlsx")

total_1200s <- merge(aspects_of_1200s_states, wars_in_1200s.nodes.stat_2, by="name")

```

```{r}

total_1000s_brokerag_reg<-total_1000s

total_1000s_brokerag_reg$win_rate <- (total_1000s_brokerag_reg$outdegree/total_1000s_brokerag_reg$totdegree)

total_1000s_brokerag_reg$loss_rate <- (total_1000s_brokerag_reg$indegree/total_1000s_brokerag_reg$totdegree)

total_1000s_brokerag_reg_binom <- total_1000s_brokerag_reg %>% mutate(more_win_or_loss = case_when(
  win_rate < 0.5 ~ 0,
    win_rate >= 0.5 ~ 1))

First_1000s_regression <- glm(more_win_or_loss~.-name-totdegree-indegree-outdegree-dc-eigen.dc-win_rate-loss_rate, total_1000s_brokerag_reg_binom, family=binomial)

First_1000s_regression


```

```{r}

set.seed(292)

total_1000s_for_regression <- total_1000s[,-c(1, 20:25)]

total_1000s_for_regression$win_rate <- (total_1000s_for_regression$outdegree/total_1000s_for_regression$totdegree)

total_1000s_for_regression$loss_rate <- (total_1000s_for_regression$indegree/total_1000s_for_regression$totdegree)

total_1000s_for_regression <- total_1000s_for_regression %>% mutate(more_win_or_loss = case_when(
  win_rate < 0.5 ~ 0,
    win_rate >= 0.5 ~ 1))

First_1000s_regression <- glm(more_win_or_loss~.-loss_rate-win_rate-totdegree-indegree-outdegree-dc-eigen.dc, total_1000s_for_regression, family=binomial)

First_1000s_regression

```

```{r}

set.seed(6738)

in_training<- sample(1:nrow(total_1000s_for_regression),  nrow(total_1000s_for_regression) * 0.7 )

training_1000s <- total_1000s_for_regression[in_training,]

test_1000s <- total_1000s_for_regression[-in_training,]

lm_1000s_binom_subset_1 <- glm(more_win_or_loss~.-loss_rate-win_rate-totdegree-indegree-outdegree-dc-eigen.dc, total_1000s_for_regression, family=binomial, subset = in_training )

logsitic_1_1000s_prob <- predict(lm_1000s_binom_subset_1, test_1000s,
type = "response")

log_preds_1<-ifelse(logsitic_1_1000s_prob >= 0.5, 1, 0)

prediction_1_logs <-mean(log_preds_1 == test_1000s$more_win_or_loss)

prediction_1_logs %>% kable()

```

```{r}

library(glmnet)
library(MASS)
```

```{r}


set.seed(246)

x_ridge <- model.matrix(more_win_or_loss ~ .-loss_rate-win_rate-totdegree-indegree-outdegree-dc-eigen.dc, total_1000s_for_regression)[, -1] 

y_ridge <- total_1000s_for_regression$more_win_or_loss

grid <- 10^seq(10, -2, length = 100)

ridge.mod <- glmnet(x_ridge, y_ridge, alpha = 0, lambda = grid)

dim(coef(ridge.mod))



```

```{r}
set.seed(729)
train_ridge <- sample(1:nrow(x_ridge), nrow(x_ridge)*0.8 ) 

test_ridge <- (-train_ridge)

y.test_ridge <- y_ridge[test_ridge]

```

```{r}
set.seed(9292)

ridge.mod <- glmnet(x_ridge[train_ridge, ], y_ridge[train_ridge], 
                    alpha = 0, lambda = grid, thresh = 1e-12)

ridge.pred <- predict(ridge.mod, s = 4, newx = x_ridge[test_ridge,])

mean((ridge.pred - y.test_ridge)^2) %>% kable()
```

```{r}
set.seed(231)
ridge.pred <- predict(ridge.mod, s = 0, newx = x_ridge[test_ridge, ], 
                      exact = T, x = x_ridge[train_ridge, ], y = y_ridge[train_ridge])

predict(ridge.mod, s = 0, exact = T, type = "coefficients", 
        x = x_ridge[train_ridge, ], y = y_ridge[train_ridge])[1:14, ]

```

```{r}

set.seed(9292)

cv.out <- cv.glmnet(x_ridge[train_ridge, ], y_ridge[train_ridge], alpha = 0) 

plot(cv.out)

```

```{r}
set.seed(9292)

bestlam <- cv.out$lambda.min

bestlam
```

```{r}
set.seed(9292)

ridge.pred <- predict(cv.out, s = bestlam, newx = x_ridge[test_ridge,])

mean((ridge.pred - y.test_ridge)^2) %>% kable()
```

```{r}

set.seed(2897)

x_lasso <- model.matrix(more_win_or_loss ~ .-loss_rate-win_rate-totdegree-indegree-outdegree-dc-eigen.dc, total_1000s_for_regression)[, -1] 

y_lasso <- total_1000s_for_regression$more_win_or_loss

grid <- 10^seq(10, -2, length = 100)

lasso.mod <- glmnet(x_lasso, y_lasso, alpha = 0, lambda = grid)

dim(coef(lasso.mod))



```

```{r}

set.seed(729)

train_lasso <- sample(1:nrow(x_ridge), nrow(x_ridge)*0.8 ) 

test_lasso <- (-train_lasso)

y.test_lasso <- y_lasso[test_lasso]
```

```{r}
set.seed(9292)

lasso.mod <- glmnet(x_lasso[train_lasso, ], y_lasso[train_lasso], 
                    alpha = 1, lambda = grid)

plot(lasso.mod)
```

```{r}

set.seed(1029)

cv.out_2 <- cv.glmnet(x_lasso[train_lasso, ], y_lasso[train_lasso], alpha = 1) 

plot(cv.out_2)

```

```{r}
set.seed(1920)

bestlam_2 <- cv.out_2$lambda.min

lasso.pred <- predict(cv.out_2, s = bestlam_2, newx = x_ridge[test_ridge,])

mean((lasso.pred - y.test_ridge)^2) %>% kable()

```

```{r}

set.seed(2739)

out <- glmnet(x_lasso[train_lasso, ], y_lasso[train_lasso], 
              alpha = 1, lambda = grid)

lasso.coef <- predict(out, type = "coefficients", s = bestlam_2)[1:14, ]

lasso.coef

```


### Community Grouping

#### Label Propagation 1000s:

The first community cluster below is done using label propagation. This results in 39 groups

```{r}
set.seed(23)
comm.lab<-label.propagation.community(wars_in_1000s.ig)
#Inspect clustering object
# igraph::groups(comm.lab)

```

```{r, echo=FALSE, warning=FALSE}

set.seed(123)

plot(comm.lab,wars_in_1000s.ig, vertex.size=0.5, edge.size=0.25, edge.arrow.size=.25,  vertex.label.dist=1.5,vertex.label.cex=0.25,asp = 0)


```

#### Walktrap 1000s:

Walktrap classification as seen below results in 19 distinct communities.

```{r}

set.seed(238)
#Run clustering algorithm: fast_greedy
wars_in_1000s.wt<-walktrap.community(wars_in_1000s.ig)

#igraph::groups(wars_in_1000s.wt)

```

Adding more steps resulted in 19 groups for both 10 and 20 steps.

```{r}
#Run & inspect clustering algorithm: 10 steps
#igraph::groups(walktrap.community(wars_in_1000s.ig, steps=10)) 
#Run & inspect clustering algorithm: 20 steps
#igraph::groups(walktrap.community(wars_in_1000s.ig ,steps=20))
#Run & inspect clustering algorithm
```

```{r, echo=FALSE}

plot(wars_in_1000s.wt, wars_in_1000s.ig, vertex.size=0.5, edge.size=0.25, edge.arrow.size=.25,  vertex.label.dist=1.5,vertex.label.cex=0.25,asp = 0)


```

### Machine Learning, Regression and Principle Components:

```{r}

total_1000s_for_PCA <- total_1000s_brokerag_reg_binom[-c(20:27)]

apply(total_1000s_for_PCA[-1], 2, mean)

```

```{r}

apply(total_1000s_for_PCA[-1], 2, var)

```

```{r}

pr.out <- prcomp(total_1000s_for_PCA[-1], scale = TRUE)

```

```{r}
names(pr.out)
```

```{r}
pr.out$center
```

```{r}

pr.out$scale

```

```{r, results=FALSE, echo=FALSE}

pr.out$rotation

```

```{r}
biplot(pr.out, scale = 0)
```

```{r}

set.seed(172)

ggbiplot(pr.out, labels =  total_1000s_for_PCA$name, labels.size  =1.5)

```


```{r}

pr.out$rotation = -pr.out$rotation 

pr.out$x = -pr.out$x

biplot(pr.out, scale = 0)

```

```{r}

pr.out$sdev

```

```{r}

pr.var <- pr.out$sdev^2

pr.var

```

```{r}

pve <- pr.var / sum(pr.var)

pve

```

```{r}

par(mfrow = c(1, 2))
plot(pve, xlab = "Principal Component",
ylab = "Proportion of Variance Explained", ylim = c(0, 1),
type = "b")

plot(cumsum(pve), xlab = "Principal Component",
ylab = "Cumulative Proportion of Variance Explained", ylim = c(0, 1), type = "b")

```

```{r}

names(total_1200s)

```

```{r}

total_1200s_brokerag_reg<-total_1200s

```

```{r}
total_1200s_brokerag_reg$win_rate <- (total_1200s_brokerag_reg$outdegree/total_1200s_brokerag_reg$totdegree)
```

```{r}
total_1200s_brokerag_reg$loss_rate <- (total_1200s_brokerag_reg$indegree/total_1200s_brokerag_reg$totdegree)
```

```{r}

total_1200s_brokerag_reg_binom <- total_1200s_brokerag_reg %>% mutate(more_win_or_loss = case_when(
  win_rate < 0.5 ~ 0,
    win_rate >= 0.5 ~ 1))

```

```{r}

total_1200s_for_PCA <- total_1200s_brokerag_reg_binom[-c(20:27)]


apply(total_1200s_for_PCA[-1], 2, mean)

```

```{r}

apply(total_1200s_for_PCA[-1], 2, var)

```

```{r}

# I cannot scale variables with 

total_1200s_for_PCA<-total_1200s_for_PCA[-c(8,10)]

```

```{r}

pr.out_2 <- prcomp(total_1200s_for_PCA[-1], scale = TRUE)

```

```{r}
names(pr.out_2)
```

```{r}
pr.out_2$center
```

```{r}

pr.out_2$scale

```

```{r, results=FALSE, echo=FALSE}
pr.out_2$rotation
```

```{r}
biplot(pr.out_2, scale = 0)
```

```{r}

pr.out_2$rotation = -pr.out_2$rotation 

pr.out_2$x = -pr.out_2$x

biplot(pr.out_2, scale = 0)

```
```{r}

set.seed(8192)

ggbiplot(pr.out_2, labels =  total_1200s_for_PCA$name, labels.size  =1.5)
```

```{r}

pr.out$sdev

```

```{r}

pr.var_2 <- pr.out_2$sdev^2

pr.var_2

```

```{r}

pve_2 <- pr.var_2 / sum(pr.var_2)

pve_2

```

```{r}

par(mfrow = c(1, 2))
plot(pve_2, xlab = "Principal Component",
ylab = "Proportion of Variance Explained", ylim = c(0, 1),
type = "b")

plot(cumsum(pve_2), xlab = "Principal Component",
ylab = "Cumulative Proportion of Variance Explained", ylim = c(0, 1), type = "b")

```

### CUG 1000s

```{r}

trans.cug<-cug.test(wars_1000s,FUN=gtrans,mode="digraph",cmode="size")
trans.cug

```

```{r}

plot(trans.cug)

```

```{r}
#t-stat between observed and simulated networks
(trans.cug$obs.stat-mean(trans.cug$rep.stat))/sd(trans.cug$rep.stat)
```

```{r}
cug.t<-function(cug.object){
  (cug.object$obs.stat-mean(cug.object$rep.stat))/sd(cug.object$rep.stat)
}
```

```{r}
#compare network transitivity to null conditional on size
trans.cug<-cug.test(wars_1000s,FUN=gtrans,mode="digraph",cmode="size", reps = 100)
trans.cug
```

```{r}
#plot vs. simulation results
plot(trans.cug)
```

```{r}
#t-stat between observed and simulated networks
cug.t(trans.cug)
```

### CUG 1100s

```{r}

trans.cug<-cug.test(wars_1100s,FUN=gtrans,mode="digraph",cmode="size")
trans.cug

```

```{r}

plot(trans.cug)

```

```{r}
#t-stat between observed and simulated networks
(trans.cug$obs.stat-mean(trans.cug$rep.stat))/sd(trans.cug$rep.stat)
```

```{r}
cug.t<-function(cug.object){
  (cug.object$obs.stat-mean(cug.object$rep.stat))/sd(cug.object$rep.stat)
}
```

```{r}
#compare network transitivity to null conditional on size
trans.cug<-cug.test(wars_1100s,FUN=gtrans,mode="digraph",cmode="size", reps = 100)
trans.cug
```

```{r}
#plot vs. simulation results
plot(trans.cug)
```

```{r}
#t-stat between observed and simulated networks
cug.t(trans.cug)
```

### CUG 1200s

```{r}

trans.cug<-cug.test(wars_1200s,FUN=gtrans,mode="digraph",cmode="size")
trans.cug

```

```{r}

plot(trans.cug)

```

```{r}
#t-stat between observed and simulated networks
(trans.cug$obs.stat-mean(trans.cug$rep.stat))/sd(trans.cug$rep.stat)
```

```{r}
cug.t<-function(cug.object){
  (cug.object$obs.stat-mean(cug.object$rep.stat))/sd(cug.object$rep.stat)
}
```

```{r}
#compare network transitivity to null conditional on size
trans.cug<-cug.test(wars_1200s,FUN=gtrans,mode="digraph",cmode="size", reps = 100)
trans.cug
```

```{r}
#plot vs. simulation results
plot(trans.cug)
```

```{r}
#t-stat between observed and simulated networks
cug.t(trans.cug)
```

### CUG Test Centralization 1000s

```{r}
#compare network degree centralization to null conditional on size
c.degree.cug <-cug.test(wars_1000s,FUN=centralization,  FUN.arg=list(FUN=degree, cmode="indegree"), mode="digraph", cmode="size") 
#plot vs simulation results
plot(c.degree.cug)
```

```{r}
#t-stat between observed and simulated networks
cug.t(c.degree.cug)
```

```{r}
#compare network betweenness centralization to null conditional on size
b.degree.cug <-cug.test(wars_1000s,FUN=centralization,  FUN.arg=list(FUN=betweenness, cmode="directed"), mode="digraph", cmode="size", reps=100) 
#plot vs simulation results
plot(b.degree.cug)
```

```{r}
#t-stat between observed and simulated networks
cug.t(b.degree.cug)
```

### CUG Test Centralization 1200s

```{r}
#compare network degree centralization to null conditional on size
c.degree.cug <-cug.test(wars_1100s,FUN=centralization,  FUN.arg=list(FUN=degree, cmode="indegree"), mode="digraph", cmode="size") 
#plot vs simulation results
plot(c.degree.cug)
```

```{r}
#t-stat between observed and simulated networks
cug.t(c.degree.cug)
```

```{r}
#compare network betweenness centralization to null conditional on size
b.degree.cug <-cug.test(wars_1100s,FUN=centralization,  FUN.arg=list(FUN=betweenness, cmode="directed"), mode="digraph", cmode="size", reps=100) 
#plot vs simulation results
plot(b.degree.cug)
```

```{r}
#t-stat between observed and simulated networks
cug.t(b.degree.cug)
```

### CUG Test Centralization 1200s

```{r}
#compare network degree centralization to null conditional on size
c.degree.cug <-cug.test(wars_1200s,FUN=centralization,  FUN.arg=list(FUN=degree, cmode="indegree"), mode="digraph", cmode="size") 
#plot vs simulation results
plot(c.degree.cug)
```

```{r}
#t-stat between observed and simulated networks
cug.t(c.degree.cug)
```

```{r}
#compare network betweenness centralization to null conditional on size
b.degree.cug <-cug.test(wars_1200s,FUN=centralization,  FUN.arg=list(FUN=betweenness, cmode="directed"), mode="digraph", cmode="size", reps=100) 
#plot vs simulation results
plot(b.degree.cug)
```

```{r}
#t-stat between observed and simulated networks
cug.t(b.degree.cug)
```

### Conditioning on Different Network Properties 1000s

```{r}
#compare network transitivity to null conditional on dyads
trans.cug<-cug.test(wars_1000s,FUN=gtrans,mode="digraph",cmode="dyad")
trans.cug
```

```{r}
#plot vs simulation results
plot(trans.cug)
```

```{r}
#t-stat between observed and simulated networks
cug.t(trans.cug)
```

```{r}
#compare network transitivity to null conditional on edges (density)
trans.cug<-cug.test(wars_1000s,FUN=gtrans,mode="digraph",cmode="edges", reps=100)
trans.cug
```

```{r}
#plot vs simulation results
plot(trans.cug)
```

```{r}
#t-stat between observed and simulated networks
cug.t(trans.cug)
```

### Conditioning on Different Network Properties 1100s

```{r}
#compare network transitivity to null conditional on dyads
trans.cug<-cug.test(wars_1100s,FUN=gtrans,mode="digraph",cmode="dyad")
trans.cug
```

```{r}
#plot vs simulation results
plot(trans.cug)
```

```{r}
#t-stat between observed and simulated networks
cug.t(trans.cug)
```

```{r}
#compare network transitivity to null conditional on edges (density)
trans.cug<-cug.test(wars_1100s,FUN=gtrans,mode="digraph",cmode="edges", reps=100)
trans.cug
```

```{r}
#plot vs simulation results
plot(trans.cug)
```

```{r}
#t-stat between observed and simulated networks
cug.t(trans.cug)
```

### Conditioning on Different Network Properties 1200s

```{r}
#compare network transitivity to null conditional on dyads
trans.cug<-cug.test(wars_1200s,FUN=gtrans,mode="digraph",cmode="dyad")
trans.cug
```

```{r}
#plot vs simulation results
plot(trans.cug)
```

```{r}
#t-stat between observed and simulated networks
cug.t(trans.cug)
```

```{r}
#compare network transitivity to null conditional on edges (density)
trans.cug<-cug.test(wars_1200s,FUN=gtrans,mode="digraph",cmode="edges", reps=100)
trans.cug
```

```{r}
#plot vs simulation results
plot(trans.cug)
```

```{r}
#t-stat between observed and simulated networks
cug.t(trans.cug)
```

```{r}
#create empty dataframe for simulations
trials<-data.frame(id=1:100, gdens=NA, gtrans=NA, cent.deg=NA, cent.bet=NA)
```

```{r}
#simulate PA networks and add stats to trials dataframe: size
for ( i in 1:100 ){ 
  pa.ig<- igraph::sample_pa(n = network.size(wars_in_1000s_network), directed=TRUE)
  pa.stat<-intergraph::asNetwork(pa.ig)
  trials$gdens<-gden(pa.stat)
  trials$gtrans[i] <- gtrans(pa.stat)
  trials$cent.deg[i] <- centralization(pa.stat, FUN=degree, cmode="indegree")
  trials$cent.bet[i] <-centralization(pa.stat, FUN=betweenness)
}
```

```{r}
summary(trials)

```

```{r}
sim.t<-function(g, trials){
  temp<-data.frame(density=c(gden(g),mean(trials$gdens),sd(trials$gdens)),
             transitivity=c(gtrans(g),mean(trials$gtrans),sd(trials$gtrans)),
             indegCent=c(centralization(g, FUN=degree, cmode="indegree"),mean(trials$cent.deg), sd(trials$cent.deg)),
             betwCent=c(centralization(g, FUN=betweenness), mean(trials$cent.bet), sd(trials$cent.bet)))
  rownames(temp)<-c("Observed","Simulated", "SD")
  temp<-data.frame(t(temp))
  temp$tvalue<-(temp$Observed-temp$Simulated)/temp$SD
  temp
}
```

```{r}
plot.sim.t<-function(g,trials){
  temp<-data.frame(net.stat=c("gtrans","cent.deg","cent.bet"), x=c(gtrans(g),centralization(g, FUN=degree, cmode="indegree"), centralization(g, FUN=betweenness)))
  trials%>%
    select(gtrans:cent.bet)%>%
    gather(key="net.stat",value = "estimate")%>%
    ggplot(aes(estimate)) +
    geom_histogram() +
    facet_wrap(net.stat ~ ., scales="free", ncol=3) +
    geom_vline(data=temp, aes(xintercept=x),
               linetype="dashed", size=1, colour="red")
}
```


```{r}
#check for differences from null

sim.t(g=wars_1000s, trials)

```

```{r}

# plot.sim.t(wars_1000s, trials)

```


